### 2024-09-23 paper/paper.qmd
Initial commit

---
title: "Crime Frequency and Population Density: Analyzing Crime Rates in Toronto Neighborhoods"
author: 
  - Yongqi Liu
thanks: "Code and data are available at: https://github.com/Cassieliu77/starter_folder-master.git."
date: 24 September 2024
date-format: long
toc: true 
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

# install.packages(c("ggplot2", "dplyr", "viridis", "sf", "patchwork", "gt"))
library(tidyverse)
library(httr)
library(opendatatoronto)
library(ggplot2)
library(knitr)
#library(dplyr)
#library(viridis)  
#library(gt)
#library(patchwork)
```

# Introduction
Urban crime trends have long been studied due to their impact on social stability and policy-making. In Toronto, crime rates vary significantly across neighborhoods, often influenced by factors such as population density and economic disparity. This paper focuses on analyzing the trends in crime frequency across different types of crimes, using data from 2014 to 2023. The objective is to determine how crime patterns have changed over time and which neighborhoods are most affected.

# Data {#sec-data}
@toronto_crime_data used in this study is sourced from @opendatatoronto and covers the period between 2014 and 2023. It includes detailed records of various crime types such as Assault, Auto Theft, Break and Enter, Robbery, Theft Over, Homicide, and Shooting & Breakenter across 158 neighborhoods in Toronto. Population data for each neighborhood is also included, enabling an analysis of the relationship between population density and crime frequency.
Following @Oliveira argument that crime increases nonlinearly with population size, particularly in more populated areas, this analysis also considers the nonlinear nature of crime frequency in Toronto neighborhoods. Crime rates per 100,000 residents were used to compare neighborhoods with different population sizes, allowing for a more accurate understanding of crime dynamics. Additionally, the dataset includes the crime rate per 100,000 residents, which is calculated using population estimates from Environics Analytics.The rate is determined by dividing the crime count by the population and multiplying by 100,000, giving a per capita measure of crime annually. It allows more accurate comparisons of crime levels across areas with different population sizes. Unlike raw crime counts, crime rates offer a more balanced comparison over time as they account for population changes in the area.
The data is analyzed in R (@citeR), using packages such as Tidyverse (@tidyverse), Httr (@httr), and ggplot2 (@ggplot2) in order to visualize the data and show important information intuitively.
@fig-rawdata shows

```{r}
#| label: fig-rawdata
#| fig-cap: the glimpse of the raw data
#| echo: false
crime_data <- read.csv(here::here("data/raw_data/raw_crime_data.csv"))
crime_data <- crime_data |> select(HOOD_ID, AREA_NAME, POPULATION_2023, ASSAULT_2023, ASSAULT_RATE_2023)
kable(head(crime_data))
```
\newpage

# Crime Rates Over the Past Decade
In order to see the overall crime rate in Toronto. The average crime rate over the 158 neighbourhood is applied to visualize @fig-rates. @fig-rates presents a depiction of how various crime types have evolved from 2014 to 2023, particularly auto theft and break-and-enter offenses. Assault remains the most prevalent crime, while homicide and robbery rates have been relatively stable. @Oliveira work suggests that theft often increases superlinearly with population size, and Toronto’s recent surge in auto theft may reflect this pattern.The sharp increase in property-related crimes, such as break-and-enters, could also indicate broader economic and social shifts that align with similar trends observed in other cities globally (@Oliveira).

The data reveals a marked increase in auto theft and break-and-enter crimes, while shooting and homicide rates have remained relatively low and stable. The rise in auto theft may be attributed to increased access to technology, while economic downturns might explain the surge in property-related crimes like break and enters.
It highlights key trends in crimes such as **Assault, Auto Theft, Bike Theft, Break and Enters, Homicide, Robbery, Shooting,** and **Theft Over**, with each category visually represented by distinct colored lines. One of the most striking observations is the consistently high rate of **Assault** (represented by the red line), which has maintained its position as the most prevalent crime throughout the period. While the rate remained relatively stable between 2014 and 2020, it experienced a notable increase starting in 2022, reflecting a potentially alarming trend that may call for increased public safety measures or changes in law enforcement focus.
In contrast, **Auto Theft** (brown line) shows one of the sharpest upward trajectories, particularly after 2020. This rise in auto-related crime could be indicative of broader societal or economic shifts, such as changes in vehicle accessibility, security measures, or even the impact of the COVID-19 pandemic, which has affected economic stability and possibly driven up theft-related crimes. The stark increase in auto theft from 2021 onward could suggest an emerging challenge for Toronto’s law enforcement and policymakers as they respond to these shifts.
Meanwhile, **Break and Enters** (green line) and **Bike Theft** (also green) exhibit moderate fluctuations over the years, with **Bike Theft** showing a noticeable increase post-2021. These fluctuations in property crimes could be influenced by a range of factors, including urban development, changes in housing density, and public awareness campaigns promoting better security practices. Other crime types like **Homicide, Robbery, Shooting,** and **Theft Over** display relatively low rates and less volatility, suggesting that these categories, while serious, have not seen significant recent surges.
The data presented in @fig-rates provides an essential foundation for discussions around urban crime in Toronto. It allows us to visualize the persistence and emergence of specific crime types over time, highlighting areas where intervention may be most necessary. For instance, the increasing rates of **Auto Theft** and **Assault** should be a priority for policymakers and law enforcement agencies. The graph also reinforces the importance of data-driven decision-making in urban planning and crime prevention, as visualizing trends over time can reveal patterns and shifts that may not be immediately apparent in static data snapshots. @fig-rates serves as a powerful tool for understanding Toronto’s crime landscape, emphasizing the need for continued monitoring and adaptive strategies in response to evolving crime dynamics.

```{r}
#| label: fig-rates
#| fig-cap: the crime rates for differnet crimes in Toronto over time
#| echo: false

# Load necessary data
crime_long <- read.csv(here::here("data/analysis_data/toronto_crime_average_rates.csv"))

# Plot the data using ggplot2
ggplot(crime_long, aes(x = Year, y = value, color = Crime_Type, group = Crime_Type)) +
  geom_line(linewidth = 1) +  # Use `linewidth` instead of `size` for lines
  geom_point(size = 2) +  # Points still use `size`
  labs(title = "Average Crime Rates in Toronto Over Time",
       x = "Year", y = "Average Crime Rate per 100,000 population",
       color = "Crime Type") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels
```
\newpage

#Results
## Toronto Population Distribution
```{r, fig.width=10, fig.height=8}
#| label: fig-Population
#| fig-cap: Population distribution in Toronto
#| echo: false
#| warning: false
#| message: false

crime_data_sf<- readRDS(here::here("data/analysis_data/cleaning_crime_data.rds"))
ggplot(crime_data_sf) +
  geom_sf(aes(fill = POPULATION_2023), color = "black", alpha = 0.9) +  
  scale_fill_gradient(low = "#d3e5ff", high = "#08306b", name = "Population") +  
  geom_text(aes(label = HOOD_ID, geometry = geometry), 
            stat = "sf_coordinates", size = 4, color = "white", check_overlap = TRUE) + 
  labs(title = "Toronto Neighbourhood Population Map",
       subtitle = "Based on 2023 Population",
       x = "Longitude", y = "Latitude") +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),  
    plot.title = element_text(size = 16, face = "bold"), 
    plot.subtitle = element_text(size = 12)  
  )
```

```{r, fig.width=10, fig.height=8}
#| label: fig-percrime
#| fig-cap: a faceted map to visualize different crimes in 2023 by neighborhood
#| echo: false
#| warning: false
#| message: false

crime_long_1 <- read_rds(here::here("data/analysis_data/cleaning_crime_count.rds"))
ggplot(crime_long_1) +
  geom_sf(aes(fill = Crime_Count), color = "black", size = 0.2, alpha = 1) +  
  scale_fill_gradientn(colors = c("white", "lightpink", "red", "darkred"),
                       values = scales::rescale(c(0, 10, 50, 100, max(crime_long_1$Crime_Count))),
                       name = "Crime Count") + 
  facet_wrap(~ Crime_Type) +
  labs(title = "Crime Rates in 2023 by Neighborhood",
       subtitle = "Visualized for Multiple Crime Types",
       fill = "Crime Count") +
  theme_minimal() +
  theme(axis.title = element_blank(),  
        panel.grid = element_blank(),   
        strip.text = element_text(size = 14, face = "bold")) 
```
\newpage
##Geographic Distribution of Crime
A geographic heatmap of Toronto’s crime shows a higher concentration of crimes in areas with higher population densities, particularly in economically disadvantaged neighborhoods. This finding shows that population size is a strong predictor of crime, with more densely populated areas experiencing higher crime rates. In Toronto, eastern neighborhoods report higher assault and robbery rates than western, more affluent areas.

## Neighborhood Analysis
The analysis of crime frequency by neighborhood over a ten-year span shows that certain neighborhoods, such as Moss Park and Kensington-Chinatown, rank among the highest in crime occurrence (@fig-rank). Conversely, neighborhoods like Kingsway South and Markland Wood have consistently remained among the safest. These trends indicate that areas with lower socioeconomic status and higher population densities are more prone to higher crime rates, reflecting similar patterns observed in other major cities.

```{r}
#| label: fig-rank
#| fig-cap: Visualize those safestest and most crime porne neighbourhood
#| echo: false
#| warning: false
#| message: false

top_10_neighborhoods_consistent <- read.csv(here::here("data/analysis_data/top_10_neighborhoods_consistent.csv"))
bottom_10_neighborhoods_consistent<- read.csv(here::here("data/analysis_data/bottom_10_neighborhoods_consistent.csv"))

ggplot(top_10_neighborhoods_consistent, aes(x = reorder(Neighborhood_ID_Name, Average_Rank), y = Average_Rank)) +
  geom_bar(stat = "identity", fill = "tomato", width = 0.7) + 
  coord_flip() +  
  labs(title = "Top 10 Crime Prone Neighborhoods (2014-2023)",
       x = "Neighborhood", 
       y = "Average Rank") +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 12, face = "bold"), 
    axis.text = element_text(size = 10),  
    axis.title = element_text(size = 12),  
    panel.grid.minor = element_blank(),  
    panel.grid.major.x = element_blank(),  
    panel.grid.major.y = element_line(color = "gray", linetype = "dashed")  
  )

ggplot(bottom_10_neighborhoods_consistent, aes(x = reorder(Neighborhood_ID_Name, -Average_Rank), y = Average_Rank)) +
  geom_bar(stat = "identity", fill = "steelblue", width = 0.7) +  
  coord_flip() + 
  labs(title = "Top 10 Safest Neighborhoods (2014-2023)",
       x = "Neighborhood", 
       y = "Average Rank") +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 12, face = "bold"),  
    axis.text = element_text(size = 10),  
    axis.title = element_text(size = 12),  
    panel.grid.minor = element_blank(),  
    panel.grid.major.x = element_blank(),  
    panel.grid.major.y = element_line(color = "gray", linetype = "dashed")  
  )
```
\newpage

# Discussion

## First discussion point {#sec-first-point}
The increasing trend in crimes like auto theft and assault calls for immediate attention from policymakers. As shown in the data, crime rates do not arise uniformly across all neighborhoods but are concentrated in specific areas. This concentration suggests that socioeconomic factors, along with population density, play a significant role in the prevalence of crime. The sharp increase in auto theft post-2020 highlights the need for enhanced security measures, while the stable yet high assault rates point to the need for sustained law enforcement efforts in vulnerable neighborhoods.

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this. 
## Second discussion point

## Third discussion point

## Weaknesses and next steps
In conclusion, the data illustrates a clear correlation between population density and crime frequency in Toronto. The most affected neighborhoods tend to be those with higher population densities and economic disadvantages. This analysis underscores the importance of using data-driven approaches to inform policy decisions, especially when it comes to resource allocation for crime prevention and community support.

\newpage
During the process, LLM is used for searching literatures.

\newpage
# References

### 2024-09-23 scripts/00-simulate_data.R

@@ -1,0 +1,84 @@
+#### Preamble ####
+# Purpose: Simulate the data
+# Author: Yongqi Liu
+# Date: 18 Sep 2024 
+# Contact: cassieliu.liu@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: Construct the sketch about the data frame roughly
+# Any other information needed? NA
+
+#### Workspace setup ####
+library(sf) 
+library(tidyverse)
+library(dplyr)
+set.seed(666)
+
+#### Simulate data ####
+n_neighborhoods <- 158
+years <- 2014:2023
+crime_types <- c("Assault", "Robbery", "Break and Enter", "Theft Over", 
+                 "Auto Theft", "Bike Theft", "Shooting", "Homicide")
+neighborhood_data <- data.frame(
+  HOOD_ID = 1:n_neighborhoods,
+  AREA_NAME = paste("Neighborhood", 1:n_neighborhoods)
+)
+head(neighborhood_data)
+
+# Simulate crime counts for each crime type, each year and neighborhood
+simulate_crime_counts <- function() {
+  sapply(crime_types, function(x) rpois(1, lambda = sample(50:200, 1)))  # Random counts between 50 and 200
+}
+crime_data <- expand.grid(
+  HOOD_ID = neighborhood_data$HOOD_ID,
+  Year = years
+) %>%
+  rowwise() %>%
+  mutate(
+    Assault = rpois(1, lambda = 100),
+    Robbery = rpois(1, lambda = 50),
+    `Break and Enter` = rpois(1, lambda = 70),
+    `Theft Over` = rpois(1, lambda = 30),
+    `Auto Theft` = rpois(1, lambda = 40),
+    `Bike Theft` = rpois(1, lambda = 20),
+    Shooting = rpois(1, lambda = 10),
+    Homicide = rpois(1, lambda = 5)
+  ) %>%
+  ungroup()
+crime_data <- crime_data %>%
+  left_join(neighborhood_data, by = "HOOD_ID")
+head(crime_data)
+
+# Simulate crime rates (per 100,000 population)
+crime_data <- crime_data %>%
+  mutate(Population = sample(5000:15000, n(), replace = TRUE)) %>%
+  mutate(across(Assault:Homicide, ~ . / Population * 100000, .names = "{.col}_Rate"))
+head(crime_data)
+
+# Create a function to simulate rankings based on crime counts
+simulate_rankings <- function(data) {
+  data %>%
[…]

### 2024-09-23 scripts/01-download_data.R

@@ -1,0 +1,31 @@
+#### Preamble ####
+# Purpose: Download the needed dataset and save it.
+# Author: Yongqi Liu
+# Date: 20 Sep 2024 
+# Contact: cassieliu.liu@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: Find the dataset called "About Neighborhood Crime Rates" on Open Data Toronto
+# Any other information needed? NA
+
+
+#### Workspace setup ####
+library(opendatatoronto)
+library(tidyverse)
+library(dplyr)
+
+#### Download data ####
+# get the package
+package <- show_package("neighbourhood-crime-rates")
+package
+# get all resources for this package
+resources <- list_package_resources("neighbourhood-crime-rates")
+# identify datastore resources; by default, Toronto Open Data sets datastore resource format to CSV for non-geospatial and GeoJSON for geospatial resources
+datastore_resources <- filter(resources, tolower(format) %in% c('csv', 'geojson'))
+# load the first datastore resource as a sample
+data <- filter(datastore_resources, row_number()==1) %>% get_resource()
+data
+
+#### Save data ####
+write_csv(data, "data/raw_data/raw_crime_data.csv") 
+
+         

### 2024-09-23 scripts/02-data_cleaning.R

@@ -1,0 +1,149 @@
+#### Preamble ####
+# Purpose: Clean the raw data and make it available for construcing graphs
+# Author: Yongqi Liu
+# Date: 20 Sep 2024 
+# Contact: cassieliu.liu@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: Have the raw data ready for processing
+
+#### Workspace setup ####
+#install.packages("sf") for handling the spatial data
+#install.packages("geojsonsf")
+library(geojsonsf)
+library(sf)
+library(tidyverse)
+library(dplyr)
+
+#### Clean data ####
+# Handling the spatial data, split the geometry column into longitude and latitude
+crime_data <- read.csv("data/raw_data/raw_crime_data.csv")
+extract_coordinates <- function(geometry_list) {
+  coordinates <- unlist(strsplit(gsub("list\\(c\\(|\\)", "", geometry_list), ","))
+  coordinates <- as.numeric(coordinates) 
+  n <- length(coordinates) / 2
+  longitudes <- coordinates[1:n]
+  latitudes <- coordinates[(n+1):(2*n)]
+  matrix(c(longitudes, latitudes), ncol = 2, byrow = FALSE)
+}
+# Apply the function to each row of the geometry column and get a list of coordinate matrices
+coordinates_list <- lapply(crime_data$geometry, extract_coordinates)
+crime_data_sf <- lapply(coordinates_list, function(coords) {
+  if (all(coords[1, ] == coords[nrow(coords), ])) {
+    st_polygon(list(coords))  
+  } else {
+    st_linestring(coords) 
+  }
+})
+crime_data$geometry <- st_sfc(crime_data_sf)
+crime_data_sf <- st_as_sf(crime_data)
+
+print(st_geometry(crime_data_sf))
+
+# Get the average crime rate in the whole Toronto for the line chart
+crime_rate_columns <- grep("RATE", names(crime_data), value = TRUE)
+average_rates <- crime_data %>%
+  summarise(across(all_of(crime_rate_columns), mean, na.rm = TRUE))
+average_rates <- cbind(AREA_NAME = "Toronto Average", average_rates)
+head(average_rates)
+crime_long <- average_rates %>%
+  pivot_longer(cols = contains("RATE"), 
+               names_to = c("Crime_Type", "Year"),  
[…]

### 2024-09-23 scripts/03-test_data.R

@@ -1,0 +1,93 @@
+#### Preamble ####
+# Purpose: Check the validity of the cleaned datasets 
+# Author: Yongqi Liu
+# Date: 21 Sep 2024 
+# Contact: cassieliu.liu@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: Obtain and have the cleaned dataset ready
+
+#### Workspace setup ####
+library(tidyverse)
+
+#### Test data ####
+# Test for Missing Values in the crime_data
+crime_data <- read.csv("data/raw_data/simulated_crime_data.csv")
+missing_values_check <- colSums(is.na(crime_data))
+print(missing_values_check)
+
+# Check for missing values in critical columns (e.g., crime counts, population)
+missing_values_critical <- crime_data_ranked %>%
+  select(Assault:Homicide, Population) %>%
+  summarise_all(~ sum(is.na(.)))
+print("Missing Values in Critical Columns:")
+print(missing_values_critical)
+# Check the structure of the dataset to confirm data types
+str(crime_data_ranked)
+
+# Specifically, check if the crime columns are numeric and the neighborhood columns are characters
+check_data_types <- sapply(crime_data_ranked, class)
+print("Data Types Check:")
+print(check_data_types)
+
+# Check that crime counts and rates are within reasonable ranges
+crime_data_range_check <- crime_data_ranked %>%
+  summarise(across(Assault:Homicide, list(min = min, max = max)),
+            across(ends_with("Rate"), list(min = min, max = max)),
+            Population_min = min(Population), Population_max = max(Population))
+
+print("Crime Counts and Rates Range Check:")
+print(crime_data_range_check)
+
+# Ensure that crime rates are non-negative
+crime_rate_negative_check <- crime_data_ranked %>%
+  summarise(across(ends_with("Rate"), ~ sum(. < 0)))
+print("Negative Crime Rates Check:")
+print(crime_rate_negative_check)
+# Check for duplicate neighborhood IDs
+duplicate_neighborhood_ids <- crime_data_ranked %>%
+  group_by(HOOD_ID) %>%
+  summarise(n = n()) %>%
+  filter(n > 1)
+
+print("Duplicate Neighborhood IDs Check:")
+print(duplicate_neighborhood_ids)
+
+# Check for duplicate neighborhood names
+duplicate_neighborhood_names <- crime_data_ranked %>%
+  group_by(AREA_NAME) %>%
[…]

### 2024-09-23 scripts/03-test_data.R
Update the test data

@@ -8,6 +8,7 @@
 
 #### Workspace setup ####
 library(tidyverse)
+library(sf)
 
 #### Test data ####
 # Test for Missing Values in the crime_data
@@ -15,79 +16,71 @@ crime_data <- read.csv("data/raw_data/simulated_crime_data.csv")
 missing_values_check <- colSums(is.na(crime_data))
 print(missing_values_check)
 
-# Check for missing values in critical columns (e.g., crime counts, population)
-missing_values_critical <- crime_data_ranked %>%
-  select(Assault:Homicide, Population) %>%
-  summarise_all(~ sum(is.na(.)))
-print("Missing Values in Critical Columns:")
-print(missing_values_critical)
-# Check the structure of the dataset to confirm data types
-str(crime_data_ranked)
+# Test for missing values
+missing_values <- crime_data %>%
+  select(Assault:Homicide, Population) %>% summarise_all(~ sum(is.na(.)))
+print(missing_values)
 
-# Specifically, check if the crime columns are numeric and the neighborhood columns are characters
-check_data_types <- sapply(crime_data_ranked, class)
-print("Data Types Check:")
+sum(is.na(crime_data))
+
+# Check the structure of the dataset to confirm data types
+str(crime_data)
+
+# Test if the crime columns are numeric and the neighborhood columns are characters
+check_data_types <- sapply(crime_data, class)
 print(check_data_types)
 
 # Check that crime counts and rates are within reasonable ranges
-crime_data_range_check <- crime_data_ranked %>%
+crime_data_range_check <- crime_data %>%
   summarise(across(Assault:Homicide, list(min = min, max = max)),
             across(ends_with("Rate"), list(min = min, max = max)),
             Population_min = min(Population), Population_max = max(Population))
-
-print("Crime Counts and Rates Range Check:")
 print(crime_data_range_check)
 
 # Ensure that crime rates are non-negative
-crime_rate_negative_check <- crime_data_ranked %>%
+crime_rate_negative_check <- crime_data %>%
   summarise(across(ends_with("Rate"), ~ sum(. < 0)))
-print("Negative Crime Rates Check:")
 print(crime_rate_negative_check)
+
 # Check for duplicate neighborhood IDs
-duplicate_neighborhood_ids <- crime_data_ranked %>%
[…]

### 2024-09-23 paper/paper.qmd
Update data analysis process, from data simulation to data cleaning. Prepare the initial draft for the paper.

@@
 thanks: "Code and data are available at: https://github.com/Cassieliu77/starter_folder-master.git."
-date: 24 September 2024
+date: 23 September 2024
 date-format: long
 toc: true 
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "This paper investigates the relationship between population density and crime rates in Toronto neighborhoods from 2014 to 2023.
+Using data from the Toronto Police Services and population estimates, the study explores how crime patterns vary across different areas of the city.This research underscores the importance of targeted crime prevention strategies and data-driven policymaking to address the specific needs of these crime-prone neighborhoods.
+Additionally, the study discusses broader societal factors influencing crime trends, including the potential impacts of economic instability and technological changes."
 format: pdf
@@
 \newpage
-During the process, LLM is used for searching literatures.
-
-\newpage
 # References

### 2024-09-23 scripts/00-simulate_data.R
Make small modifications on the simulation script's annotations.

@@ -64,14 +64,14 @@ simulate_rankings <- function(data) {
 crime_data_ranked <- simulate_rankings(crime_data)
 head(crime_data_ranked)
 
-# Simulate random geometry (latitude and longitude) for neighborhoods
+# Simulate geometry data (latitude and longitude) for neighborhoods
 simulate_geometry <- function(n) {
   latitudes <- runif(n, min = 43.6, max = 43.85)  # Toronto's latitude range
   longitudes <- runif(n, min = -79.6, max = -79.1)  # Toronto's longitude range
   st_as_sf(data.frame(
     lat = latitudes, 
     lon = longitudes
-  ), coords = c("lon", "lat"), crs = 4326)  # Create spatial points
+  ), coords = c("lon", "lat"), crs = 4326)
 }
 neighborhood_geometries <- simulate_geometry(n_neighborhoods)
 

### 2024-09-23 scripts/01-download_data.R
Download the data from the website OpeDataToronto

@@ -1,12 +1,10 @@
 #### Preamble ####
-# Purpose: Download the needed dataset and save it.
+# Purpose: Download the needed dataset
 # Author: Yongqi Liu
 # Date: 20 Sep 2024 
 # Contact: cassieliu.liu@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: Find the dataset called "About Neighborhood Crime Rates" on Open Data Toronto
-# Any other information needed? NA
-
 
 #### Workspace setup ####
 library(opendatatoronto)

### 2024-09-23 scripts/02-data_cleaning.R
Change the cleaning data name to a briefer one.

@@ -144,6 +144,6 @@ saveRDS(crime_data_sf, "data/analysis_data/cleaning_crime_data.rds")
 write_csv(crime_data_sf, "data/analysis_data/cleaning_crime_data.csv")
 write_csv(crime_long, "data/analysis_data/toronto_crime_average_rates.csv")
 write_rds(crime_long_1, "data/analysis_data/cleaning_crime_count.rds")
-write_csv(top_10_neighborhoods_consistent, "data/analysis_data/top_10_neighborhoods_consistent.csv")
-write_csv(bottom_10_neighborhoods_consistent, "data/analysis_data/bottom_10_neighborhoods_consistent.csv")
+write_csv(top_10_neighborhoods_consistent, "data/analysis_data/top_10_neighborhoods.csv")
+write_csv(bottom_10_neighborhoods_consistent, "data/analysis_data/bottom_10_neighborhoods.csv")
 

### 2024-09-23 scripts/01-download_data.R
Remove some unnecessary annotations.

@@ -17,7 +17,6 @@ package <- show_package("neighbourhood-crime-rates")
 package
 # get all resources for this package
 resources <- list_package_resources("neighbourhood-crime-rates")
-# identify datastore resources; by default, Toronto Open Data sets datastore resource format to CSV for non-geospatial and GeoJSON for geospatial resources
 datastore_resources <- filter(resources, tolower(format) %in% c('csv', 'geojson'))
 # load the first datastore resource as a sample
 data <- filter(datastore_resources, row_number()==1) %>% get_resource()

### 2024-09-24 paper/paper.qmd
Update the paper draft.

--- 2024-09-23 paper/paper.qmd
@@
 ---
-title: "Crime Frequency and Population Density: Analyzing Crime Rates in Toronto Neighborhoods"
+title: "Exploring the Impact of Population Density on Crime Distribution in Toronto Neighborhoods (2014–2023)"
 author: 
   - Yongqi Liu
-thanks: "Code and data are available at: https://github.com/Cassieliu77/starter_folder-master.git."
+thanks: "Code and data are available at:https://github.com/Cassieliu77/Crime-Frequency-and-Population-Density.git."
 date: 24 September 2024
@@
 toc: true 
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "This paper examines the relationship between population density and crime frequency in Toronto neighborhoods from 2014 to 2023, using data from Toronto Police Services.
+It analyzes how crime patterns vary across the city and identifies the 10 safest neighborhoods.
+The study highlights the need for targeted, data-driven crime prevention strategies and offers policy recommendations for law enforcement and public officials.
+It also suggests future research directions to further explore crime dynamics in urban areas."
 format: pdf
@@
 
-# install.packages(c("ggplot2", "dplyr", "viridis", "sf", "patchwork", "gt"))
+# install.packages(c("ggplot2", "dplyr", "viridis", "sf", "gt"))
 library(tidyverse)
@@
 library(knitr)
-#library(dplyr)
-#library(viridis)  
-#library(gt)
-#library(patchwork)
+library(dplyr)
+library(viridis)  
+library(gt)
 ```
@@
 The objective is to determine how crime patterns have changed over time and which neighborhoods are most affected.
+The result shows a icnreasing trend in crime rates.
+All the analysis in the paper is conducted in @citeR.
+(Need to write more!)
 
 # Data {#sec-data}
-@toronto_crime_data used in this study is sourced from @opendatatoronto and covers the period between 2014 and 2023.
-It includes detailed records of various crime types such as Assault, Auto Theft, Break and Enter, Robbery, Theft Over, Homicide, and Shooting & Breakenter across 158 neighborhoods in Toronto.
-Population data for each neighborhood is also included, enabling an analysis of the relationship between population density and crime frequency.
-Following @Oliveira argument that crime increases nonlinearly with population size, particularly in more populated areas, this analysis also considers the nonlinear nature of crime frequency in Toronto neighborhoods.
-Crime rates per 100,000 residents were used to compare neighborhoods with different population sizes, allowing for a more accurate understanding of crime dynamics.
-Additionally, the dataset includes the crime rate per 100,000 residents, which is calculated using population estimates from Environics Analytics.The rate is determined by dividing the crime count by the population and multiplying by 100,000, giving a per capita measure of crime annually.
-It allows more accurate comparisons of crime levels across areas with different population sizes.
+## Raw Data
+The Neighborhood Crime Rates dataset used in this study comes from @toronto_crime_data, which is downloaded on @opendatatoronto and covers the period from 2014 to 2023.
+It records various crime types such as Assault, Auto Theft, Break and Enter, Robbery, Theft Over, Homicide, Shooting and so on.
+The data focuses on 158 neighborhoods in Toronto and provides crime rates per 100,000 population.
+Crime rate here is measured using population estimates from Environics Analytics, which is in line with the standard definition by Statistics Canada.
+This metric allows for fairer comparisons between neighborhoods with different population sizes.
 Unlike raw crime counts, crime rates offer a more balanced comparison over time as they account for population changes in the area.
-The data is analyzed in R (@citeR), using packages such as Tidyverse (@tidyverse), Httr (@httr), and ggplot2 (@ggplot2) in order to visualize the data and show important information intuitively.
-@fig-rawdata shows
+This explains why I do not choose to use Police Annual Statistical Report - Reported Crime as my dataset.
+It only contains the reported crime count for a certain year and a certain division, which lack a standard way to compare the crime frequency among different neigborhoods.
+Besides, lack of geometry attribute makes it hard to see the distribution of crimes in the subsequent sections of the paper.
+In @toronto_crime_data, population for 2023 year in each neighborhood is also included, enabling the analysis of the relationship between population density and crime frequency.
+@fig-rawdata gives us a glimpse about how the dataset looks.
+Note that this dataset incorporates data for the new structure of the 158 City of Toronto neighborhoods, enabling a more comprehensive geographic analysis, and it does not contain occurrences of crime which were deemed “unfounded”.
+Each neighborhood is identified by their area name and a unique hood ID.
+Geometry data in the dataset is useful to build up the Toronto city map to observe population and crime distribution.
+The data is analyzed in R (@citeR), using packages such as Tidyverse (@tidyverse), Dplyr(@dplyr), Geojsonsf(@geojsonsf), Httr (@httr), Knitr(@knitr) and ggplot2 (@ggplot2) in order to visualize the data and communicate information intuitively.
 
@@
 crime_data <- read.csv(here::here("data/raw_data/raw_crime_data.csv"))
-crime_data <- crime_data |> select(HOOD_ID, AREA_NAME, POPULATION_2023, ASSAULT_2023, ASSAULT_RATE_2023)
+crime_data <- crime_data |> select(HOOD_ID,POPULATION_2023, ASSAULT_2023, ASSAULT_RATE_2023)
 kable(head(crime_data))
 ```
+
+## New Variable Constructure
+Average Crime Rate in the whole Toronto is constructed to see the crime trend in the city. It is computed based on the average of 2014-2023 year data. 
+Geometry column is transformed to the spatial data to help us make map visualizations in the Result sections.
+Average rank for each neighborhood is constructed based on the crime count for each year.
+Firstly, using the total count of crimes each year to rank 158 neighborhoods, and each neighborhood is able to get a rank for each year.
+Then, take average of it to get the average rank.The average rank of 2014-2023 is constructed based on the crime count for each year.
+The higher the crime count, the smaller the rank number.
+The summary for the top and bottom 10 ranked neighborhoods can be found in @fig-summary1 and @fig-summary2.
+
+```{r}
+#| label: fig-summary1
+#| fig-cap: Summary Table for Top10 Tanked Neighborhoods
+#| echo: false
+#| warning: false
+#| message: false
+
+top_10_neighborhoods <- read.csv(here::here("data/analysis_data/top_10_neighborhoods.csv"))
+top_summary<-top_10_neighborhoods |> select(Neighborhood_ID_Name, Average_Rank)
+
+kable(top_summary)
+```
+
+```{r}
+#| label: fig-summary2
+#| fig-cap: Summary Rable for Bottom 10 Ranked Neighborhoods
+#| echo: false
+#| warning: false
+#| message: false
+
+bottom_10_neighborhoods<- read.csv(here::here("data/analysis_data/bottom_10_neighborhoods.csv"))
+bottom_summary<-bottom_10_neighborhoods |> select(Neighborhood_ID_Name, Average_Rank)
+kable(bottom_summary)
+```
+
 \newpage
-
-# Crime Rates Over the Past Decade
+# Results
+## Crime Rates Over the Past Decade
 In order to see the overall crime rate in Toronto.
-The average crime rate over the 158 neighbourhood is applied to visualize @fig-rates.
-@fig-rates presents a depiction of how various crime types have evolved from 2014 to 2023, particularly auto theft and break-and-enter offenses.
+The average crime rate over the 158 neighbourhood is applied to visualize @fig-rates, which presents a depiction of how various crime types have evolved from 2014 to 2023, particularly auto theft and break-and-enter offenses.
 Assault remains the most prevalent crime, while homicide and robbery rates have been relatively stable.
-@Oliveira work suggests that theft often increases superlinearly with population size, and Toronto’s recent surge in auto theft may reflect this pattern.The sharp increase in property-related crimes, such as break-and-enters, could also indicate broader economic and social shifts that align with similar trends observed in other cities globally (@Oliveira).
+@Oliveira work suggests that theft often increases superlinearly with population size, and Toronto’s recent surge in auto theft may reflect this pattern.The sharp increase in property-related crimes, such as break-and-enters, could also indicate broader economic and social shifts that align with similar trends observed in other cities globally @Oliveira.
 
@@
 
-#Results
 ## Toronto Population Distribution
+@fig-Population shows the population distribution in Toronto, and help us to identify whether these crime frequency area is accordance with those high population density neighborhoods.
 ```{r, fig.width=10, fig.height=8}
@@
 ```
+\newpage
+
+## Geographic Distribution of Crime
+Based on @fig-Population, @fig-percrime represents a higher concentration of crimes in areas with higher population densities, particularly in economically disadvantaged neighborhoods.
+Those gray areas mean no crime was reported.
+From the map showing **Homicide**, it This finding shows that population size is a strong predictor of crime, with more densely populated areas experiencing higher crime rates.
+In Toronto, eastern neighborhoods report higher assault and robbery rates than western, more affluent areas.
 
@@
 #| label: fig-percrime
-#| fig-cap: a faceted map to visualize different crimes in 2023 by neighborhood
+#| fig-cap: 2023 Crime Distribution in Toronto by Neighborhoods
 #| echo: false
@@
   facet_wrap(~ Crime_Type) +
-  labs(title = "Crime Rates in 2023 by Neighborhood",
+  labs(title = "Crime Count in 2023 by Neighborhoods",
        subtitle = "Visualized for Multiple Crime Types",
@@
 \newpage
-##Geographic Distribution of Crime
-A geographic heatmap of Toronto’s crime shows a higher concentration of crimes in areas with higher population densities, particularly in economically disadvantaged neighborhoods.
-This finding shows that population size is a strong predictor of crime, with more densely populated areas experiencing higher crime rates.
-In Toronto, eastern neighborhoods report higher assault and robbery rates than western, more affluent areas.
 
 ## Neighborhood Analysis
-The analysis of crime frequency by neighborhood over a ten-year span shows that certain neighborhoods, such as Moss Park and Kensington-Chinatown, rank among the highest in crime occurrence (@fig-rank).
-Conversely, neighborhoods like Kingsway South and Markland Wood have consistently remained among the safest.
-These trends indicate that areas with lower socioeconomic status and higher population densities are more prone to higher crime rates, reflecting similar patterns observed in other major cities.
-
 ```{r}
-#| label: fig-rank
-#| fig-cap: Visualize those safestest and most crime porne neighbourhood
+#| label: fig-rank-crime
+#| fig-cap: Top10 Ranked Crime Prone Neigborhoods Based on Crime Count
 #| echo: false
@@
 
-top_10_neighborhoods_consistent <- read.csv(here::here("data/analysis_data/top_10_neighborhoods_consistent.csv"))
-bottom_10_neighborhoods_consistent<- read.csv(here::here("data/analysis_data/bottom_10_neighborhoods_consistent.csv"))
-
-ggplot(top_10_neighborhoods_consistent, aes(x = reorder(Neighborhood_ID_Name, Average_Rank), y = Average_Rank)) +
+ggplot(top_10_neighborhoods, aes(x = reorder(Neighborhood_ID_Name, Average_Rank), y = Average_Rank)) +
   geom_bar(stat = "identity", fill = "tomato", width = 0.7) + 
   coord_flip() +  
-  labs(title = "Top 10 Crime Prone Neighborhoods (2014-2023)",
+  labs(title = "Top 10 Crime Prone Neighborhoods ",
        x = "Neighborhood", 
-       y = "Average Rank") +
+       y = "Average Rank (2014-2023)") +
   theme_minimal() +
@@
     plot.title = element_text(size = 12, face = "bold"), 
-    axis.text = element_text(size = 10),  
-    axis.title = element_text(size = 12),  
-    panel.grid.minor = element_blank(),  
-    panel.grid.major.x = element_blank(),  
-    panel.grid.major.y = element_line(color = "gray", linetype = "dashed")  
-  )
-
-ggplot(bottom_10_neighborhoods_consistent, aes(x = reorder(Neighborhood_ID_Name, -Average_Rank), y = Average_Rank)) +
-  geom_bar(stat = "identity", fill = "steelblue", width = 0.7) +  
-  coord_flip() + 
-  labs(title = "Top 10 Safest Neighborhoods (2014-2023)",
-       x = "Neighborhood", 
-       y = "Average Rank") +
-  theme_minimal() +
-  theme(
-    plot.title = element_text(size = 12, face = "bold"),  
-    axis.text = element_text(size = 10),  
+    axis.text = element_text(size = 8),  
     axis.title = element_text(size = 12),  
@@
 ```
+
+```{r}
+#| label: fig-rank-safe
+#| fig-cap: Top10 Ranked Safest Neigborhoods Based on Crime Count
+#| echo: false
+#| warning: false
+#| message: false
+ggplot(bottom_10_neighborhoods, aes(x = reorder(Neighborhood_ID_Name, -Average_Rank), y = Average_Rank)) +
+  geom_bar(stat = "identity", fill = "steelblue", width = 0.7) +  
+  coord_flip() + 
+  labs(title = "Top 10 Safest Neighborhoods",
+       x = "Neighborhood", 
+       y = "Average Rank(2014-2023)") +
+  theme_minimal() +
+  theme(
+    plot.title = element_text(size = 12, face = "bold"),  
+    axis.text = element_text(size = 8),  
+    axis.title = element_text(size = 12),  
+    panel.grid.minor = element_blank(),  
+    panel.grid.major.x = element_blank(),  
+    panel.grid.major.y = element_line(color = "gray", linetype = "dashed")  
+  )
+```
+In @fig-rank-crime, neighborhoods such as West Humber-Clairville, York University Heights, and Wellington Place have consistently ranked among the top 10 crime-prone areas over the past decade.
+These neighborhoods are characterized by higher population densities and socioeconomic challenges, making them more vulnerable to crime.
+The increasing trend in auto theft and assault in these neighborhoods underscores the need for targeted interventions and enhanced security measures.
+In contrast,@fig-rank-safe neighborhoods like Maple Leaf and Lambton Baby Point have consistently ranked as some of the safest areas in Toronto.
+These neighborhoods tend to have lower population densities and are more affluent, with better access to resources and community support.
+The stability of crime rates in these areas suggests that socioeconomic factors play a critical role in maintaining low crime levels.
 \newpage
@@
 
-## First discussion point {#sec-first-point}
+## Summary of Findings
+The study investigates the relationship between population density and crime frequency across Toronto neighborhoods from 2014 to 2023.
+The data reveals clear trends, particularly with crimes such as auto theft and assault, which are more prevalent in densely populated neighborhoods.
+The findings highlight how crime rates are not uniformly distributed but instead cluster in certain areas, often driven by socioeconomic factors.
+
+##  Population Density as a Key Driver of Crime
 The increasing trend in crimes like auto theft and assault calls for immediate attention from policymakers.
@@
 The sharp increase in auto theft post-2020 highlights the need for enhanced security measures, while the stable yet high assault rates point to the need for sustained law enforcement efforts in vulnerable neighborhoods.
+The data clearly shows a significant correlation between population density and crime frequency in Toronto neighborhoods.
+Densely populated areas such as Moss Park, St.
+Lawrence-East Bayfront, and Kensington-Chinatown consistently rank among the highest in terms of crime rates.
+This supports existing criminological theories, such as Routine Activity Theory and Social Disorganization Theory, which suggest that increased population density leads to more opportunities for crime due to greater anonymity, social fragmentation, and limited community cohesion.
+The results from this study confirm that neighborhoods with more people living in close proximity are more likely to experience higher crime rates, especially property crimes like auto theft and break-and-enters.
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this. 
-## Second discussion point
+The relationship between population density and crime suggests that urban planners and local governments should consider designing interventions that specifically address high-density environments.
+Strategies such as better lighting, increased public surveillance (e.g., CCTV), and more frequent police patrols can act as deterrents in areas where crime tends to cluster.
+Additionally, promoting community cohesion through events, neighborhood associations, and resident engagement initiatives can reduce the social disorganization that often accompanies densely populated urban settings.
 
-## Third discussion point
+## Socioeconomic Disparities and Crime Hotspots
+In addition to population density, socioeconomic factors significantly influence crime patterns.
+Neighborhoods with higher levels of economic disadvantage, such as Moss Park and York University Heights, experience higher crime rates.
+In contrast, affluent neighborhoods like Kingsway South report much lower crime rates.
+This suggests that addressing socioeconomic disparities is critical to reducing crime.
+Targeted social interventions, such as improving access to education, healthcare, and employment opportunities, could alleviate some of the underlying factors contributing to crime.
+Socioeconomic disparities shape crime dynamics by contributing to social strain and increasing the motivation for economic crimes such as theft, robbery, and property damage.
+Policymakers should focus on alleviating economic inequality through social programs that address poverty, unemployment, and access to education.
+For example, neighborhood revitalization programs that improve housing conditions, increase job opportunities, and provide better access to education and healthcare could reduce the social conditions that lead to higher crime rates.
 
-## Weaknesses and next steps
-In conclusion, the data illustrates a clear correlation between population density and crime frequency in Toronto.
-The most affected neighborhoods tend to be those with higher population densities and economic disadvantages.
-This analysis underscores the importance of using data-driven approaches to inform policy decisions, especially when it comes to resource allocation for crime prevention and community support.
+##  Targeted Crime Prevention Strategies
+The geographic concentration of crime in specific neighborhoods has significant implications for law enforcement and public policy.
+Policymakers and law enforcement agencies must recognize that crime prevention cannot be approached with a one-size-fits-all strategy.
+Instead, targeted interventions are necessary to address the unique challenges faced by high-risk neighborhoods.
+he data reinforces the idea that crime is not uniformly distributed across Toronto neighborhoods but is concentrated in specific high-risk areas.
+This suggests that a “one-size-fits-all” approach to crime prevention would be ineffective.
+Instead, law enforcement agencies should focus their efforts on areas where crime is most prevalent, such as Moss Park and Kensington-Chinatown, by deploying more officers and allocating more resources to these neighborhoods.
 
-\newpage
-During the process, LLM is used for searching literatures.
+At the same time, the study highlights the importance of community-based crime prevention strategies.
+Engaging local residents in neighborhood watch programs, providing education on crime prevention techniques, and fostering collaboration between police and community organizations can empower communities to take ownership of their safety.
+This approach not only reduces crime rates but also builds trust between law enforcement and the communities they serve, which is essential for long-term success in crime reduction.
+
+For neighborhoods like Moss Park and St.
+Lawrence-East Bayfront, which consistently rank among the highest neighborhodos in crimes, a multifaceted approach is required.
+This might include increased policing and community-based programs aimed at addressing the root causes of crime, such as poverty, unemployment, and lack of access to education and healthcare.
+Research suggests that addressing these underlying issues can have a significant impact on reducing crime rates, particularly in economically disadvantaged areas.
+
+Furthermore, the rise in auto theft and break-and-enter incidents post-2020 suggests that law enforcement must adapt to changing crime patterns.
+Technological advancements, such as the increased use of vehicle tracking systems and smart security devices, could help curb the rising rates of property crime.
+In addition, public awareness campaigns aimed at promoting better security practices, particularly in high-risk neighborhoods, may also prove effective in preventing crime.
+
+##  Limitations and Future Research
+While this study represents the relationship between population density and crime frequency in Toronto, several limitations must be acknowledged.
+Firstly, the analysis relies on reported crime data, which may not fully capture the extent of criminal activity, particularly for underreported crimes such as domestic violence and cybercrime.
+
+Secondly, the study focuses primarily on the relationship between population density and crime rates, without accounting for other factors that may also influence crime, such as the presence of law enforcement, urban infrastructure, and even social events.
+Future research could incorporate these variables to provide a more nuanced understanding of crime dynamics in urban areas.
+
+Finally, the analysis spans a relatively short period (2014–2023), which may not capture long-term trends or shifts in crime patterns.
+Expanding the dataset to include earlier years or incorporating data from other cities could offer a broader perspective on crime dynamics in Toronto and beyond.
+Future research should expand on these findings by incorporating additional data sources to capture a broader range of criminal activity.
+It would also be useful to explore how different types of urban infrastructure, such as public transportation or lighting, influence crime rates.
+Finally, comparative studies with other major cities could provide a more comprehensive understanding of urban crime dynamics, enabling better-targeted interventions for crime prevention.
 

### 2024-09-26 paper/paper.qmd
Update LLM usage
Delete unused file.

@@
 The objective is to determine how crime patterns have changed over time and which neighborhoods are most affected.
-The result shows a icnreasing trend in crime rates.
-All the analysis in the paper is conducted in @citeR.
+The result shows a increasing trend in crime rates.
+All the analysis in the paper is conducted in @citeR, @openai2023chatgpt is used in this research paper to help simulate data.
 (Need to write more!)
@@
 In Toronto, eastern neighborhoods report higher assault and robbery rates than western, more affluent areas.
+```{r,fig.width=10, fig.height=8}
+#| label: fig-crimeandpopulation
+#| fig-cap: 2023 Crime Trend with Population
+#| echo: false
+#| warning: false
+#| message: false
+crime_long_1 <- read_rds(here::here("data/analysis_data/cleaning_crime_count.rds"))
+
+ggplot(crime_long_1, aes(x = Crime_Count, y = POPULATION_2023)) +
+  geom_point(aes(color = Crime_Type), size = 2, alpha = 0.8) +  # Adjust point size and transparency for better visibility
+  geom_smooth(method = "lm", se = FALSE, color = "black", size = 1) +  # Keep linear trendline with original method
+  scale_color_manual(values = c("lightpink", "red", "grey", "blue", "orange", "skyblue", "brown", "yellow")) +  # Customize colors for crime types
+  scale_x_continuous(limits = c(0, 200), breaks = seq(0, 200, by = 50)) +  # Set horizontal axis limits from 0 to 200 with breaks
+  facet_wrap(~ Crime_Type) +  # Split the scatterplot by crime type
+  labs(title = "Crime Count vs. Population by Neighborhoods",
+       subtitle = "Visualized for Multiple Crime Types",
+       x = "Crime Count",  
+       y = "Population",
+       color = "Crime Type") +
+  theme_minimal(base_size = 15) +  # Increase base size for text readability
+  theme(axis.title = element_text(size = 14),  
+        panel.grid.major = element_line(color = "gray90"),  # Add subtle grid lines for clarity
+        panel.grid.minor = element_blank(),   
+        strip.text = element_text(size = 16, face = "bold"),  # Emphasize facet labels
+        legend.position = "bottom",  # Move legend to the bottom for a cleaner layout
+        legend.title = element_text(face = "bold"),
+        plot.title = element_text(face = "bold", hjust = 0.5),  # Center align the title
+        plot.subtitle = element_text(hjust = 0.5))  # Center align the subtitle
+```
 

### 2024-09-26 scripts/02-data_cleaning.R
Delete unused csv file

@@ -61,9 +61,10 @@ crime_long <- average_rates %>%
 head(crime_long)
 
 # Clean the dataset for the combination of maps for per type of crime
+# Transform and clean crime_long_1
 crime_long_1 <- crime_data_sf %>%
   pivot_longer(cols = c(ASSAULT_2023, ROBBERY_2023, BREAKENTER_2023, THEFTOVER_2023, AUTOTHEFT_2023,
-                        BIKETHEFT_2023,SHOOTING_2023,HOMICIDE_2023), 
+                        BIKETHEFT_2023, SHOOTING_2023, HOMICIDE_2023), 
                names_to = "Crime_Type", 
                values_to = "Crime_Count") %>%
   mutate(Crime_Type = case_when(
@@ -75,8 +76,10 @@ crime_long_1 <- crime_data_sf %>%
     Crime_Type == "BIKETHEFT_2023" ~ "Bike Theft",
     Crime_Type == "HOMICIDE_2023" ~ "Homicide",
     Crime_Type == "SHOOTING_2023" ~ "Shooting",
-    TRUE ~ Crime_Type 
-  ))
+    TRUE ~ Crime_Type
+  )) %>%
+  drop_na(Crime_Count)  # Remove rows where Crime_Count is NA
+# Check the result
 head(crime_long_1)
 
 #Get the crime count for different neighbourhood in different years
@@ -144,6 +147,7 @@ saveRDS(crime_data_sf, "data/analysis_data/cleaning_crime_data.rds")
 write_csv(crime_data_sf, "data/analysis_data/cleaning_crime_data.csv")
 write_csv(crime_long, "data/analysis_data/toronto_crime_average_rates.csv")
 write_rds(crime_long_1, "data/analysis_data/cleaning_crime_count.rds")
+write_csv(crime_long_1, "data/analysis_data/cleaning_crime_count.csv")
 write_csv(top_10_neighborhoods_consistent, "data/analysis_data/top_10_neighborhoods.csv")
 write_csv(bottom_10_neighborhoods_consistent, "data/analysis_data/bottom_10_neighborhoods.csv")
 

### 2024-09-26 paper/paper.qmd
Update paper.

@@
 thanks: "Code and data are available at:https://github.com/Cassieliu77/Crime-Frequency-and-Population-Density.git."
-date: 24 September 2024
+date: 26 September 2024
 date-format: long
@@
 
-# install.packages(c("ggplot2", "dplyr", "viridis", "sf", "gt"))
+# install.packages(c("ggplot2", "dplyr", "viridis", "sf", "gt","lintr"))
 library(tidyverse)
@@
 library(viridis)
+library(sf)
 library(gt)
+library(lintr)
 ```
@@
 # Introduction
-Urban crime trends have long been studied due to their impact on social stability and policy-making.
-In Toronto, crime rates vary significantly across neighborhoods, often influenced by factors such as population density and economic disparity.
-This paper focuses on analyzing the trends in crime frequency across different types of crimes, using data from 2014 to 2023.
-The objective is to determine how crime patterns have changed over time and which neighborhoods are most affected.
-The result shows a increasing trend in crime rates.
-All the analysis in the paper is conducted in @citeR, @openai2023chatgpt is used in this research paper to help simulate data.
-(Need to write more!)
+Urban crime has long been a critical focus for researchers and policymakers due to its significant impact on societal well-being and public safety.
+In Toronto, crime rates differ considerably across neighborhoods, often influenced by population density and socioeconomic factors.
+This study examines the relationship between population density and crime rates in Toronto neighborhoods from 2014 to 2023.
+By analyzing data from Toronto Police Services, we aim to understand how crime patterns are shaped by population distribution and to provide data-driven recommendations for more targeted crime prevention strategies.
+Through this analysis, we hope to shed light on the dynamics of urban crime in Toronto and contribute to more effective policies for mitigating crime in high-density areas.
+All of the analyses in the paper is conducted in @citeR.
 
@@
 ## Raw Data
-The Neighborhood Crime Rates dataset used in this study comes from @toronto_crime_data, which is downloaded on @opendatatoronto and covers the period from 2014 to 2023.
-It records various crime types such as Assault, Auto Theft, Break and Enter, Robbery, Theft Over, Homicide, Shooting and so on.
-The data focuses on 158 neighborhoods in Toronto and provides crime rates per 100,000 population.
-Crime rate here is measured using population estimates from Environics Analytics, which is in line with the standard definition by Statistics Canada.
-This metric allows for fairer comparisons between neighborhoods with different population sizes.
+The Neighborhood Crime Rates dataset, obtained from @toronto_crime_data, spans from 2014 to 2023 and covers 158 Toronto neighborhoods.
+The dataset includes various crime types such as Assault, Auto Theft, Break and Enter, Robbery, and Homicide, expressed as rates per 100,000 people.
+Crime counts for per crime type per year are also shown in the data.
 Unlike raw crime counts, crime rates offer a more balanced comparison over time as they account for population changes in the area.
 This explains why I do not choose to use Police Annual Statistical Report - Reported Crime as my dataset.
-It only contains the reported crime count for a certain year and a certain division, which lack a standard way to compare the crime frequency among different neigborhoods.
-Besides, lack of geometry attribute makes it hard to see the distribution of crimes in the subsequent sections of the paper.
-In @toronto_crime_data, population for 2023 year in each neighborhood is also included, enabling the analysis of the relationship between population density and crime frequency.
+It only contains the reported crime count for a certain year and a certain division, which lack a standard way to compare the crime frequency among different neighborhoods.
+Besides, lack of spatial attribute makes it hard to see the geography distribution of crimes in the subsequent sections of the paper.
+In @toronto_crime_data, another used variable population for 2023 year in each neighborhood is also included, enabling the analysis of the relationship between population density and crime frequency.
+Additionally, new variables, such as the average crime rate across Toronto from 2014 to 2023 and average neighborhood rankings based on crime counts, were constructed to analyze long-term trends and identify high-risk areas.
+As for the measurement, crime rate is measured using population estimates from Environics Analytics, which is in line with the standard definition by Statistics Canada.
+The dataset records crimes that got Toronto Police intervention.
+This metric allows for fairer comparisons between neighborhoods with different population sizes.
+Also note that this dataset incorporates data for the new structure of the 158 City of Toronto neighborhoods, enabling a more comprehensive geographic analysis, and it does not contain occurrences of crime which were deemed “unfounded”.
 @fig-rawdata gives us a glimpse about how the dataset looks. 
-Note that this dataset incorporates data for the new structure of the 158 City of Toronto neighborhoods, enabling a more comprehensive geographic analysis, and it does not contain occurrences of crime which were deemed “unfounded”.
-Each neighborhood is identified by their area name and a unique hood ID.
-Geometry data in the dataset is useful to build up the Toronto city map to observe population and crime distribution.
-The data is analyzed in R (@citeR), using packages such as Tidyverse (@tidyverse), Dplyr(@dplyr), Geojsonsf(@geojsonsf), Httr (@httr), Knitr(@knitr) and ggplot2 (@ggplot2) in order to visualize the data and communicate information intuitively.
+Each neighborhood is identified by their area name and a unique hood ID. Geometry data in the dataset is used to build up the Toronto city map to observe population density and crime distribution.
+The data is analyzed in R (@citeR), using packages Tidyverse (@tidyverse), Dplyr(@dplyr), Geojsonsf(@geojsonsf), Httr (@httr), Knitr(@knitr) and ggplot2 (@ggplot2) in order to visualize the data and communicate information intuitively.
 
@@
 #| label: fig-rawdata
-#| fig-cap: the glimpse of the raw data
+#| fig-cap: A Lookup for the Raw Data
 #| echo: false
@@
 ## New Variable Constructure
+Crime_Type and Crime_Count are constructed based on 2023 data. Crime_
 Average Crime Rate in the whole Toronto is constructed to see the crime trend in the city. It is computed based on the average of 2014-2023 year data. 
@@
 The summary for the top and bottom 10 ranked neighborhoods can be found in @fig-summary1 and @fig-summary2.
+```{r}
+#| label: fig-summary0
+#| fig-cap: Summary Table for the Crime Count
+#| echo: false
+#| warning: false
+#| message: false
+
+crime_count <- readRDS(here::here("data/analysis_data/cleaning_crime_count.rds"))
+crime_count_summary <- crime_count %>%
+  st_drop_geometry() %>%  # Remove the geometry column
+  select(HOOD_ID, AREA_NAME, Crime_Type, Crime_Count)
+# Display the first 10 rows of the selected summary
+kable(head(crime_count_summary, 10))
+```
 
@@
 @fig-rates serves as a powerful tool for understanding Toronto’s crime landscape, emphasizing the need for continued monitoring and adaptive strategies in response to evolving crime dynamics.
+It is noticeable that 2020 is a turning point, when the pandemic begins. The crime rate experience a dramatic increase, expecially for Assault, increasing from 600 to 800 per 100,000 population.
 
@@
 ## Geographic Distribution of Crime
-Based on @fig-Population, @fig-percrime represents a higher concentration of crimes in areas with higher population densities, particularly in economically disadvantaged neighborhoods.
+Based on @fig-Population, @fig-crimeandpopulation represents a higher concentration of crimes in areas with higher population densities, particularly in economically disadvantaged neighborhoods.
+The black trend line here shows an Upward divergence in scatterplot for differnt crime types.
+
+Such effects arise from nonlinear population effects that persist in rates due to the linear assumption.
 Those gray areas mean no crime was reported.
@@
 
-```{r, fig.width=10, fig.height=8}
-#| label: fig-percrime
-#| fig-cap: 2023 Crime Distribution in Toronto by Neighborhoods
+From @fig-crimeandpopulation, these scatter plots represent upward trend between Crime Count and Population is shown.
+It seems that the lack of cases in **Shooting** and **Homicide** makes most dots overlap together.
+So, the map for each crime is created respectively to help us see the distribution of these 2 particular crime types.
+
+```{r,fig.width=10, fig.height=8}
+#| label: fig-homicide
+#| fig-cap: 2023 Homicide Distribution
 #| echo: false
@@
 
-crime_long_1 <- read_rds(here::here("data/analysis_data/cleaning_crime_count.rds"))
-ggplot(crime_long_1) +
-  geom_sf(aes(fill = Crime_Count), color = "black", size = 0.2, alpha = 1) +  
-  scale_fill_gradientn(colors = c("white", "lightpink", "red", "darkred"),
-                       values = scales::rescale(c(0, 10, 50, 100, max(crime_long_1$Crime_Count))),
-                       name = "Crime Count") + 
-  facet_wrap(~ Crime_Type) +
-  labs(title = "Crime Count in 2023 by Neighborhoods",
-       subtitle = "Visualized for Multiple Crime Types",
-       fill = "Crime Count") +
+homicide_data <- crime_long_1 %>%
+  filter(Crime_Type == "Homicide")
+# Check the filtered data
+#head(homicide_data)
+ggplot(crime_data_sf) +
+  geom_sf(aes(fill = POPULATION_2023), color = "black", alpha = 0.8) +  # Population density map with slight transparency
+  scale_fill_gradient(low = "#d3e5ff", high = "#08306b", name = "Population") +  # Softer gradient for better visuals
+  geom_jitter(data = homicide_data, aes(geometry = geometry, size = Crime_Count),  # Use geom_jitter to avoid overlapping dots
+              color = "red", stat = "sf_coordinates", alpha = 0.6, width = 0.002, height = 0.002) +  # Add jitter and adjust transparency
+  scale_size_continuous(name = "Homicide Count", range = c(2, 10)) +  # Dot sizes for clearer representation of crime count
+  labs(title = "Homicide Count and Population Density",
+       subtitle = "By Neighborhoods",
+       x = "Longitude", y = "Latitude") +
   theme_minimal() +
-  theme(axis.title = element_blank(),  
-        panel.grid = element_blank(),   
-        strip.text = element_text(size = 14, face = "bold")) 
+  theme(
+    panel.grid.major = element_blank(),  
+    panel.grid.minor = element_blank(),  
+    plot.title = element_text(size = 18, face = "bold", hjust = 0.5),  
+    plot.subtitle = element_text(size = 14, hjust = 0.5),  
+    legend.position = "bottom", 
+    legend.title = element_text(face = "bold", size = 12),  
+    legend.text = element_text(size = 10)  
+  ) +
+  guides(size = guide_legend(override.aes = list(alpha = 1))) 
 ```
+```{r,fig.width=10, fig.height=8}
+#| label: fig-shooting
+#| fig-cap: Shooting Count Distribution in 2023
+#| echo: false
+#| warning: false
+#| message: false
+
+# Filter for Shooting data
+shooting_data <- crime_long_1 %>%
+  filter(Crime_Type == "Shooting")
+
+# Create a population density map with shooting crime count
+ggplot(crime_data_sf) +
+  geom_sf(aes(fill = POPULATION_2023), color = "black", alpha = 0.9) +  # Population density map
+  scale_fill_gradient(low = "#d3e5ff", high = "#08306b", name = "Population") +  # Color gradient for population density
+  geom_jitter(data = shooting_data, aes(geometry = geometry, size = Crime_Count),  # Add jitter to avoid overlap
+              color = "red", stat = "sf_coordinates", alpha = 0.6, width = 0.002, height = 0.002) +  # Jitter points
+  scale_size_continuous(name = "Shooting Count", range = c(1, 10), breaks = c(5, 10, 15, 20)) +  # Set dot size range and breaks
+  labs(title = "Shooting Count and Population Density",
+       subtitle = "By Neighborhood",
+       x = "Longitude", y = "Latitude") +
+  theme_minimal() +
+  theme(
+    panel.grid = element_blank(),  # Remove grid
+    plot.title = element_text(size = 18, face = "bold"), 
+    plot.subtitle = element_text(size = 14),
+    legend.position = "bottom",  # Move legend to bottom
+    legend.title = element_text(face = "bold")
+  ) +
+  guides(size = guide_legend(override.aes = list(alpha = 1)))  # Adjust legend for point size
+```
+
 \newpage
@@
 ## Neighborhood Analysis
+After looking at the trend of crimes in the whole Toronto, it is more important to take a look into those small neighborhoods.
+Top10 safest and Top10 Crime Prone neighborhoods are shown in the following content.
+It give police and security department to take different measures to differnt neighborhoods.
+
 ```{r}
@@
 ## Summary of Findings
-The study investigates the relationship between population density and crime frequency across Toronto neighborhoods from 2014 to 2023.
-The data reveals clear trends, particularly with crimes such as auto theft and assault, which are more prevalent in densely populated neighborhoods.
-The findings highlight how crime rates are not uniformly distributed but instead cluster in certain areas, often driven by socioeconomic factors.
+The analysis revealed significant variations in crime rates over time, particularly in property-related offenses such as Auto Theft and Break and Enter.
+Assault remains the most common crime, while Homicide and Robbery rates have remained stable.
+The data shows an upward trend in crime rates after 2020, particularly in Assault and Auto Theft, suggesting a link between population density and increased criminal activity, especially post-pandemic.
 
@@
 Lawrence-East Bayfront, and Kensington-Chinatown consistently rank among the highest in terms of crime rates.
-This supports existing criminological theories, such as Routine Activity Theory and Social Disorganization Theory, which suggest that increased population density leads to more opportunities for crime due to greater anonymity, social fragmentation, and limited community cohesion.
-The results from this study confirm that neighborhoods with more people living in close proximity are more likely to experience higher crime rates, especially property crimes like auto theft and break-and-enters.
-
+The findings support criminological theories suggesting that increased population density creates more opportunities for crime due to social fragmentation and anonymity.
+Urban planners and policymakers must consider interventions such as improved public surveillance and community engagement initiatives to reduce crime in high-density areas.
 The relationship between population density and crime suggests that urban planners and local governments should consider designing interventions that specifically address high-density environments.
@@
 ## Socioeconomic Disparities and Crime Hotspots
-In addition to population density, socioeconomic factors significantly influence crime patterns.
-Neighborhoods with higher levels of economic disadvantage, such as Moss Park and York University Heights, experience higher crime rates.
-In contrast, affluent neighborhoods like Kingsway South report much lower crime rates.
-This suggests that addressing socioeconomic disparities is critical to reducing crime.
-Targeted social interventions, such as improving access to education, healthcare, and employment opportunities, could alleviate some of the underlying factors contributing to crime.
-Socioeconomic disparities shape crime dynamics by contributing to social strain and increasing the motivation for economic crimes such as theft, robbery, and property damage.
-Policymakers should focus on alleviating economic inequality through social programs that address poverty, unemployment, and access to education.
-For example, neighborhood revitalization programs that improve housing conditions, increase job opportunities, and provide better access to education and healthcare could reduce the social conditions that lead to higher crime rates.
+Socioeconomic factors, alongside population density, play a significant role in shaping crime patterns.
+Neighborhoods with higher levels of economic disadvantage, such as Moss Park and York University Heights, tend to have elevated crime rates, while affluent areas like Kingsway South report significantly lower crime levels.
+This suggests that addressing socioeconomic disparities is essential for reducing crime.
+Targeted social interventions, such as improving access to education, healthcare, and employment opportunities, could help mitigate the underlying causes of crime.
+Crime rates also vary between urban and rural areas.
+Densely populated urban areas typically experience higher crime rates due to greater social anonymity, economic disparities, and more opportunities for criminal activity.
+Conversely, remote rural areas often face challenges related to limited police resources, complicating crime prevention and law enforcement efforts.
+Socioeconomic inequality drives crime dynamics by creating social strain and increasing the likelihood of economic crimes such as theft, robbery, and property damage.
+Policymakers should prioritize reducing economic inequality through programs that address poverty, unemployment, and lack of access to education.
+Neighborhood revitalization initiatives that improve housing, create job opportunities, and enhance access to essential services like education and healthcare could help address the root causes of crime and reduce its prevalence in disadvantaged areas.
 
@@
 Firstly, the analysis relies on reported crime data, which may not fully capture the extent of criminal activity, particularly for underreported crimes such as domestic violence and cybercrime.
+@Oliveira's findings indicate that the assumption of linear crime growth is inaccurate.
+In over half of the analyzed data sets, we observed evidence of nonlinear crime growth, meaning that crime tends to increase with population size at a rate different from per capita expectations.
+This nonlinearity introduces a population effect that impacts crime rates and alters city rankings.
+We showed that ranking cities based on crime rates alone yields significantly different results compared to rankings that account for population size.
+So, the further
 
@@
 Expanding the dataset to include earlier years or incorporating data from other cities could offer a broader perspective on crime dynamics in Toronto and beyond.
+
+Besides, @Wang mentions the digital era has ushered in a new type of criminal activity: cybercrime and telecom crime.
+The Canadian Centre for Cyber Security (CCCS) reports a consistent rise in cybercrime incidents, which amounted to 74,073 police-reported cases as of recent data from Statista @Wang.
+This kind of crime count is missing in our dataset.While obtaining a definitive national cybercrime rate is challenging due to underreporting, the CCCS highlights phishing scams, malware attacks, and ransomware as significant threats.
+
 Future research should expand on these findings by incorporating additional data sources to capture a broader range of criminal activity.
@@
 \newpage
+# Acknowledgements
+We would like to express our appreciation to the City of Toronto and the Open Data Portal for granting access to the dataset via the Open Data Toronto package [@opendatatoronto].
+Besides, R codes from @citerohan are used to provide tables in this paper.The provision of high-quality, open-source data is essential for enabling precise research.
+Thanks to @openai2023chatgpt, which is used in this research paper.Code written in the scripts was checked and styled with lintr [@citelintr] and styler [@citestyler].
+
 # References

### 2024-09-26 scripts/02-data_cleaning.R
Delete unused csv file.

@@ -147,7 +147,6 @@ saveRDS(crime_data_sf, "data/analysis_data/cleaning_crime_data.rds")
 write_csv(crime_data_sf, "data/analysis_data/cleaning_crime_data.csv")
 write_csv(crime_long, "data/analysis_data/toronto_crime_average_rates.csv")
 write_rds(crime_long_1, "data/analysis_data/cleaning_crime_count.rds")
-write_csv(crime_long_1, "data/analysis_data/cleaning_crime_count.csv")
 write_csv(top_10_neighborhoods_consistent, "data/analysis_data/top_10_neighborhoods.csv")
 write_csv(bottom_10_neighborhoods_consistent, "data/analysis_data/bottom_10_neighborhoods.csv")
 

### 2024-09-26 paper/paper.qmd
Update the paper draft.

@@
 toc: true 
-abstract: "This paper examines the relationship between population density and crime frequency in Toronto neighborhoods from 2014 to 2023, using data from Toronto Police Services.
-It analyzes how crime patterns vary across the city and identifies the 10 safest neighborhoods.
-The study highlights the need for targeted, data-driven crime prevention strategies and offers policy recommendations for law enforcement and public officials.
-It also suggests future research directions to further explore crime dynamics in urban areas."
+abstract: "This study examines the relationship between population density and crime distribution in Toronto neighborhoods from 2014 to 2023.
+Using data from Toronto Police Services, the paper analyzes how crime patterns vary across the city and identifies key neighborhoods where crime occurance is concentrated.By focusing on the types of crime most influenced by urban density, this study offers data-driven recommendations for targeted crime prevention strategies and policy interventions.
+The findings highlight the need for more nuanced urban planning and law enforcement efforts to address the underlying causes of crime in high-density areas."
 format: pdf
@@
 library(dplyr)
-library(viridis)
+#library(viridis)
 library(sf)
@@
 # Introduction
-Urban crime has long been a critical focus for researchers and policymakers due to its significant impact on societal well-being and public safety.
-In Toronto, crime rates differ considerably across neighborhoods, often influenced by population density and socioeconomic factors.
-This study examines the relationship between population density and crime rates in Toronto neighborhoods from 2014 to 2023.
-By analyzing data from Toronto Police Services, we aim to understand how crime patterns are shaped by population distribution and to provide data-driven recommendations for more targeted crime prevention strategies.
+Urban crime has long been a focus for researchers and policymakers due to its profound impact on public safety and quality of life.
+In a rapidly growing metropolis like Toronto, population density plays a significant role in shaping the spatial distribution of crime.
+This study explores how population density correlates with crime frequency across Toronto neighborhoods between 2014 and 2023.
+Additionally, the research examines the influence of socioeconomic factors in high-density areas, aiming to identify neighborhoods that require targeted crime prevention strategies.
 Through this analysis, we hope to shed light on the dynamics of urban crime in Toronto and contribute to more effective policies for mitigating crime in high-density areas.
-All of the analyses in the paper is conducted in @citeR.
+All of the analyses in the paper are conducted in @citeR.
 
 # Data {#sec-data}
-## Raw Data
-The Neighborhood Crime Rates dataset, obtained from @toronto_crime_data, spans from 2014 to 2023 and covers 158 Toronto neighborhoods.
-The dataset includes various crime types such as Assault, Auto Theft, Break and Enter, Robbery, and Homicide, expressed as rates per 100,000 people.
-Crime counts for per crime type per year are also shown in the data.
-Unlike raw crime counts, crime rates offer a more balanced comparison over time as they account for population changes in the area.
-This explains why I do not choose to use Police Annual Statistical Report - Reported Crime as my dataset.
-It only contains the reported crime count for a certain year and a certain division, which lack a standard way to compare the crime frequency among different neighborhoods.
-Besides, lack of spatial attribute makes it hard to see the geography distribution of crimes in the subsequent sections of the paper.
-In @toronto_crime_data, another used variable population for 2023 year in each neighborhood is also included, enabling the analysis of the relationship between population density and crime frequency.
-Additionally, new variables, such as the average crime rate across Toronto from 2014 to 2023 and average neighborhood rankings based on crime counts, were constructed to analyze long-term trends and identify high-risk areas.
-As for the measurement, crime rate is measured using population estimates from Environics Analytics, which is in line with the standard definition by Statistics Canada.
+## Data Source
+The Neighborhood Crime Rates dataset, obtained from @opendatatoronto, spans from 2014 to 2023 and covers 158 Toronto neighborhoods.It is designed to offer communities insights into public safety and awareness.
+The data provided to the Toronto Police Service by reporting parties is preliminary and may not have undergone full verification at the time of publication.
+The dataset includes various crime types such as Assault, Auto Theft, Robbery, Break and Enter, Bike Theft, Shooting, and Homicide.
+According to the standard definition by Statistics Canada, the crime rate is calculated as the number of crimes per 100,000 people each year.
+Besides, crime counts per crime type per year are also shown in the data.
+In @opendatatoronto, another variable of interest *POPULATION_2023* refers to the population in the 2023 year in each neighborhood, enabling the analysis of the relationship between population density and crime frequency.
+Geometry column is transformed into spatial data to build up the Toronto city map to observe population density and crime distribution.
+
+As for the measurement, the crime rate is measured using population estimates from Environics Analytics, which is in line with the standard definition by Statistics Canada.
 The dataset records crimes that got Toronto Police intervention.
 This metric allows for fairer comparisons between neighborhoods with different population sizes.
-Also note that this dataset incorporates data for the new structure of the 158 City of Toronto neighborhoods, enabling a more comprehensive geographic analysis, and it does not contain occurrences of crime which were deemed “unfounded”.
-@fig-rawdata gives us a glimpse about how the dataset looks. 
-Each neighborhood is identified by their area name and a unique hood ID. Geometry data in the dataset is used to build up the Toronto city map to observe population density and crime distribution.
-The data is analyzed in R (@citeR), using packages Tidyverse (@tidyverse), Dplyr(@dplyr), Geojsonsf(@geojsonsf), Httr (@httr), Knitr(@knitr) and ggplot2 (@ggplot2) in order to visualize the data and communicate information intuitively.
+Also note that this dataset incorporates data for the new structure of the 158 City of Toronto neighborhoods, enabling a more comprehensive geographic analysis, and it does not contain occurrences of crime that were deemed “unfounded”.
+Fields have been updated to reflect the new structure of the 158 City of Toronto neighborhoods.Population figures only account for the resident population in each region and do not include temporary populations like commuters or business patrons.
+To protect the privacy of those involved, the locations of crime incidents have been intentionally shifted to the nearest road intersection node.
+As such, all location data should be regarded as approximate, and users are cautioned not to interpret these locations as linked to any specific address or individual.
+
+Crime rates offer a more balanced comparison over time as they account for population changes in the area and give us a fairer way to see the crime trends in the past decade, which explains why a similar dataset called "Police Annual Statistical Report - Reported Crimes" is not be chosen as my dataset, which only contains the reported crime count for a certain year and a certain division, which lack a standard way to show the crime trends and spatial attribute to see the geography distribution.
+
+@fig-rawdata gives us a glimpse of how the dataset looks. Each neighborhood is identified by its *AREA_NAME* and a unique *HOOD_ID*.
 
@@
 #| echo: false
+
 crime_data <- read.csv(here::here("data/raw_data/raw_crime_data.csv"))
-crime_data <- crime_data |> select(HOOD_ID,POPULATION_2023, ASSAULT_2023, ASSAULT_RATE_2023)
+crime_data <- crime_data |> select(HOOD_ID, POPULATION_2023, ASSAULT_2023, ASSAULT_RATE_2023)
 kable(head(crime_data))
@@
 ## New Variable Constructure
-Crime_Type and Crime_Count are constructed based on 2023 data. Crime_
-Average Crime Rate in the whole Toronto is constructed to see the crime trend in the city. It is computed based on the average of 2014-2023 year data. 
-Geometry column is transformed to the spatial data to help us make map visualizations in the Result sections.
-Average rank for each neighborhood is constructed based on the crime count for each year.
-Firstly, using the total count of crimes each year to rank 158 neighborhoods, and each neighborhood is able to get a rank for each year.
-Then, take average of it to get the average rank.The average rank of 2014-2023 is constructed based on the crime count for each year.
-The higher the crime count, the smaller the rank number.
-The summary for the top and bottom 10 ranked neighborhoods can be found in @fig-summary1 and @fig-summary2.
+The average crime rates represent the mean crime rate across 158 neighborhoods from 2014 to 2023, and were calculated to analyze long-term trends and highlight crime types that have seen significant increases.
+The variables *Crime_Type* and *Crime_Count* are based on 2023 data (@fig-summary1).
+
+Each neighborhood’s average rank was determined by ranking all 158 neighborhoods based on their total annual crime count.
+A rank is assigned for each year, and the average rank over the past decade (2014–2023) is calculated by averaging these yearly ranks.
+A lower rank indicates a higher crime count.
+Summaries of the top 10 and bottom 10 ranked neighborhoods are provided in @fig-summary2 and @fig-summary3.
 ```{r}
-#| label: fig-summary0
-#| fig-cap: Summary Table for the Crime Count
+#| label: fig-summary1
+#| fig-cap: Summary Table for the 2023 Crime Count
 #| echo: false
@@
 crime_count_summary <- crime_count %>%
-  st_drop_geometry() %>%  # Remove the geometry column
+  st_drop_geometry() %>%  
   select(HOOD_ID, AREA_NAME, Crime_Type, Crime_Count)
-# Display the first 10 rows of the selected summary
 kable(head(crime_count_summary, 10))
@@
 ```{r}
-#| label: fig-summary1
-#| fig-cap: Summary Table for Top10 Tanked Neighborhoods
+#| label: fig-summary2
+#| fig-cap: Summary Table for Top10 Ranked Neighborhoods
 #| echo: false
@@
 top_summary<-top_10_neighborhoods |> select(Neighborhood_ID_Name, Average_Rank)
-
 kable(top_summary)
@@
 ```{r}
-#| label: fig-summary2
-#| fig-cap: Summary Rable for Bottom 10 Ranked Neighborhoods
+#| label: fig-summary3
+#| fig-cap: Summary Table for Bottom 10 Ranked Neighborhoods
 #| echo: false
@@
 ## Crime Rates Over the Past Decade
-In order to see the overall crime rate in Toronto.
-The average crime rate over the 158 neighbourhood is applied to visualize @fig-rates, which presents a depiction of how various crime types have evolved from 2014 to 2023, particularly auto theft and break-and-enter offenses.
-Assault remains the most prevalent crime, while homicide and robbery rates have been relatively stable.
-@Oliveira work suggests that theft often increases superlinearly with population size, and Toronto’s recent surge in auto theft may reflect this pattern.The sharp increase in property-related crimes, such as break-and-enters, could also indicate broader economic and social shifts that align with similar trends observed in other cities globally @Oliveira.
+@fig-rates uses the variable *Average_Crime_Rate* (@fig-summary4) to show the crime rate trend between 2014 and 2023 and demonstrate the evolution of different crime types in the city of Toronto.
+It can be seen that the visualization reveals certain patterns in some specific crime types.
 
-The data reveals a marked increase in auto theft and break-and-enter crimes, while shooting and homicide rates have remained relatively low and stable.
-The rise in auto theft may be attributed to increased access to technology, while economic downturns might explain the surge in property-related crimes like break and enters.
-It highlights key trends in crimes such as **Assault, Auto Theft, Bike Theft, Break and Enters, Homicide, Robbery, Shooting,** and **Theft Over**, with each category visually represented by distinct colored lines.
-One of the most striking observations is the consistently high rate of **Assault** (represented by the red line), which has maintained its position as the most prevalent crime throughout the period.
-While the rate remained relatively stable between 2014 and 2020, it experienced a notable increase starting in 2022, reflecting a potentially alarming trend that may call for increased public safety measures or changes in law enforcement focus.
-In contrast, **Auto Theft** (brown line) shows one of the sharpest upward trajectories, particularly after 2020.
-This rise in auto-related crime could be indicative of broader societal or economic shifts, such as changes in vehicle accessibility, security measures, or even the impact of the COVID-19 pandemic, which has affected economic stability and possibly driven up theft-related crimes.
-The stark increase in auto theft from 2021 onward could suggest an emerging challenge for Toronto’s law enforcement and policymakers as they respond to these shifts.
-Meanwhile, **Break and Enters** (green line) and **Bike Theft** (also green) exhibit moderate fluctuations over the years, with **Bike Theft** showing a noticeable increase post-2021.
-These fluctuations in property crimes could be influenced by a range of factors, including urban development, changes in housing density, and public awareness campaigns promoting better security practices.
-Other crime types like **Homicide, Robbery, Shooting,** and **Theft Over** display relatively low rates and less volatility, suggesting that these categories, while serious, have not seen significant recent surges.
-The data presented in @fig-rates provides an essential foundation for discussions around urban crime in Toronto.
-It allows us to visualize the persistence and emergence of specific crime types over time, highlighting areas where intervention may be most necessary.
-For instance, the increasing rates of **Auto Theft** and **Assault** should be a priority for policymakers and law enforcement agencies.
-The graph also reinforces the importance of data-driven decision-making in urban planning and crime prevention, as visualizing trends over time can reveal patterns and shifts that may not be immediately apparent in static data snapshots.
-@fig-rates serves as a powerful tool for understanding Toronto’s crime landscape, emphasizing the need for continued monitoring and adaptive strategies in response to evolving crime dynamics.
-It is noticeable that 2020 is a turning point, when the pandemic begins. The crime rate experience a dramatic increase, expecially for Assault, increasing from 600 to 800 per 100,000 population.
+```{r}
+#| label: fig-summary4
+#| fig-cap: Average Crime Rate in Toronto
+#| echo: false
+#| warning: false
+#| message: false
+
+average_crime_rate <- read.csv(here::here("data/analysis_data/toronto_crime_average_rates.csv"))
+#top_summary<-top_10_neighborhoods |> select(Neighborhood_ID_Name, Average_Rank)
+average_crime_rate |>
+  slice(1:10) |>
+  kable()
+```
+
+@fig-rates reveals a significant upward trend in **Auto Theft** and **Assault**, particularly after 2020.
+The crime rate for **Assault** consistently remained the highest across all crime types, starting at around 600 per 100,000 population in 2014 and reaching a peak of approximately 800 per 100,000 population in 2023.
+This consistent rise suggests that assault has been a persistent issue in Toronto, with an even sharper increase following the onset of the COVID-19 pandemic.**Break and Enter**,**Robbery** and **Theft Over** all experienced increases after 2021.
+This rise suggests a sharp increase in private property related crimes, possibly influenced by socioeconomic factors or changes in law enforcement or technology.
+In contrast, **Homicide** and **Shooting** had relatively lower crime rates and displayed more stability over time, with little significant change compared to **Assault** and **Auto Theft**.
 
@@
 #| label: fig-rates
-#| fig-cap: the crime rates for differnet crimes in Toronto over time
+#| fig-cap: Average Crime Rates in Toronto Over Time
 #| echo: false
+#| warning: false
+#| message: false
 
-# Load necessary data
 crime_long <- read.csv(here::here("data/analysis_data/toronto_crime_average_rates.csv"))
-
-# Plot the data using ggplot2
-ggplot(crime_long, aes(x = Year, y = value, color = Crime_Type, group = Crime_Type)) +
-  geom_line(linewidth = 1) +  # Use `linewidth` instead of `size` for lines
-  geom_point(size = 2) +  # Points still use `size`
+ggplot(crime_long, aes(x = Year, y = Average_Crime_Rate, color = Crime_Type, group = Crime_Type)) +
+  geom_line(linewidth = 1) +  
+  geom_point(size = 2) +  
   labs(title = "Average Crime Rates in Toronto Over Time",
@@
   theme_minimal() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))  
 ```
@@
 
-## Toronto Population Distribution
-@fig-Population shows the population distribution in Toronto, and help us to identify whether these crime frequency area is accordance with those high population density neighborhoods.
-```{r, fig.width=10, fig.height=8}
+## Toronto Population Density
+@fig-Population shows the population density in Toronto, constructed by using variable *POPULATION_2023* and geometry data.
+Higher saturation represents higher population density in 2023.
+Neighborhoods with lighter shading have smaller populations, such as neighborhood 12 and 13.
+Notably, neighborhoods 1 and 95 exhibit the darkest saturation, indicating the highest population density.
+In the following sections, we will explore whether these high-density neighborhoods correspond to areas that are more prone to crime.
+
+```{r, fig.width=12, fig.height=10}
 #| label: fig-Population
-#| fig-cap: Population distribution in Toronto
+#| fig-cap: Population Density in Toronto
 #| echo: false
@@
     plot.title = element_text(size = 16, face = "bold"), 
-    plot.subtitle = element_text(size = 12)  
-  )
+    plot.subtitle = element_text(size = 12)  )
 ```
@@
 
-## Geographic Distribution of Crime
-Based on @fig-Population, @fig-crimeandpopulation represents a higher concentration of crimes in areas with higher population densities, particularly in economically disadvantaged neighborhoods.
-The black trend line here shows an Upward divergence in scatterplot for differnt crime types.
+## Relationship between Crime Count and Population 
+The scatterplots in @fig-crimeandpopulation illustrate the relationship between crime count and population for various crime types across Toronto neighborhoods in 2023.
+Each crime type is represented by a separate plot, and a linear trendline has been added to demonstrate the relationship between crime count and population size.
 
-Such effects arise from nonlinear population effects that persist in rates due to the linear assumption.
-Those gray areas mean no crime was reported.
-From the map showing **Homicide**, it This finding shows that population size is a strong predictor of crime, with more densely populated areas experiencing higher crime rates.
-In Toronto, eastern neighborhoods report higher assault and robbery rates than western, more affluent areas.
-```{r,fig.width=10, fig.height=8}
+```{r,fig.width=8, fig.height=6}
 #| label: fig-crimeandpopulation
-#| fig-cap: 2023 Crime Trend with Population
+#| fig-cap: Crime Count with Population in 2023
 #| echo: false
@@
 #| message: false
+
 crime_long_1 <- read_rds(here::here("data/analysis_data/cleaning_crime_count.rds"))
-
 ggplot(crime_long_1, aes(x = Crime_Count, y = POPULATION_2023)) +
-  geom_point(aes(color = Crime_Type), size = 2, alpha = 0.8) +  # Adjust point size and transparency for better visibility
-  geom_smooth(method = "lm", se = FALSE, color = "black", size = 1) +  # Keep linear trendline with original method
-  scale_color_manual(values = c("lightpink", "red", "grey", "blue", "orange", "skyblue", "brown", "yellow")) +  # Customize colors for crime types
-  scale_x_continuous(limits = c(0, 200), breaks = seq(0, 200, by = 50)) +  # Set horizontal axis limits from 0 to 200 with breaks
-  facet_wrap(~ Crime_Type) +  # Split the scatterplot by crime type
+  geom_point(aes(color = Crime_Type), size = 2, alpha = 0.8) +  
+  geom_smooth(method = "lm", se = FALSE, color = "black", size = 1) +  
+  scale_color_manual(values = c("lightpink", "red", "grey", "blue", "orange", "skyblue", "brown", "yellow")) +   
+  scale_x_continuous(limits = c(0, 200), breaks = seq(0, 200, by = 50)) +  
+  facet_wrap(~ Crime_Type) +  
   labs(title = "Crime Count vs. Population by Neighborhoods",
-       subtitle = "Visualized for Multiple Crime Types",
        x = "Crime Count",  
@@
        color = "Crime Type") +
-  theme_minimal(base_size = 15) +  # Increase base size for text readability
-  theme(axis.title = element_text(size = 14),  
-        panel.grid.major = element_line(color = "gray90"),  # Add subtle grid lines for clarity
+  theme_minimal(base_size = 13) +  
+  theme(axis.title = element_text(size = 12),  
+        panel.grid.major = element_line(color = "gray90"), 
         panel.grid.minor = element_blank(),   
-        strip.text = element_text(size = 16, face = "bold"),  # Emphasize facet labels
-        legend.position = "bottom",  # Move legend to the bottom for a cleaner layout
+        strip.text = element_text(size = 14, face = "bold"),  
+        legend.position = "bottom",  
         legend.title = element_text(face = "bold"),
-        plot.title = element_text(face = "bold", hjust = 0.5),  # Center align the title
-        plot.subtitle = element_text(hjust = 0.5))  # Center align the subtitle
+        plot.title = element_text(face = "bold", hjust = 0.5), 
+        plot.subtitle = element_text(hjust = 0.5))  
 ```
+**Assault** and **Auto Theft** exhibit strong positive correlations with population, as indicated by the upward slope of the trendlines.
+This suggests that as the population of a neighborhood increases, the number of assaults and auto theft incidents tends to rise as well.
+**Break and Enter** shows a similar trend, with higher crime counts in neighborhoods with larger populations.
 
+In contrast, crimes such as **Homicide** and **Shooting** display much smaller crime counts overall, and their trendlines suggest weaker correlations with population size.
+These crimes are less frequent, and the relationship between population and crime count is less pronounced compared to property crimes.
+
+**Robbery** and **Theft Over** also show a positive relationship with population, although the correlation is not as strong as it is for Assault or Auto Theft.
+**Bike Theft**, while showing an upward trend, appears more dispersed, indicating a weaker correlation between population size and the number of incidents.
+
+Overall, the graph highlights that property crimes, such as **Assault**, **Auto Theft**, and **Break and Enter**, are more likely to increase with population density, whereas violent crimes like **Homicide** and **Shooting** remain relatively low regardless of neighborhood size.
 From @fig-crimeandpopulation, these scatter plots represent upward trend between Crime Count and Population is shown.
@@
 So, the map for each crime is created respectively to help us see the distribution of these 2 particular crime types.
+
+As for the Homicide and Shooting,.............
 
@@
     legend.position = "bottom", 
-    legend.title = element_text(face = "bold", size = 12),  
-    legend.text = element_text(size = 10)  
+    legend.title = element_text(face = "bold", size = 6),  
+    legend.text = element_text(size = 6)  
   ) +
   guides(size = guide_legend(override.aes = list(alpha = 1)))
+
 ```
+
 ```{r,fig.width=10, fig.height=8}
@@
 Top10 safest and Top10 Crime Prone neighborhoods are shown in the following content.
-It give police and security department to take different measures to differnt neighborhoods.
+It give police and security department to take different measures to differnt neighborhoods.Top 1 crime prone neighborhoods is West Humber-Clairville.
+Based on the map, it lcoates
 
@@
 Targeted social interventions, such as improving access to education, healthcare, and employment opportunities, could help mitigate the underlying causes of crime.
+@fig-rates shows several crime types, notably Assault and Auto Theft, experienced noticeable increases after 2020, possibly indicating societal shifts, such as the impacts of the pandemic on public safety and economic conditions.
 Crime rates also vary between urban and rural areas.
@@
 ##  Targeted Crime Prevention Strategies
+@Wang's spatial analyses support previous scholarly findings that crime is not randomly dispersed across areas, but is instead concentrated in neighborhoods that exhibit specific characteristics.
 The geographic concentration of crime in specific neighborhoods has significant implications for law enforcement and public policy.
@@
 This kind of crime count is missing in our dataset.While obtaining a definitive national cybercrime rate is challenging due to underreporting, the CCCS highlights phishing scams, malware attacks, and ransomware as significant threats.
+Besides, as for differnt crime types, there may be differnt potential factors influecning the crime occurance.
 
@@
 # Acknowledgements
+All data is analyzed in R (@citeR), using packages Tidyverse (@tidyverse), Dplyr(@dplyr), Geojsonsf(@geojsonsf), Httr (@httr), Knitr(@knitr) and ggplot2 (@ggplot2), sf(@sf), gt(@gt), lintr(@lintr) in order to visualize the data and communicate information intuitively.
+
 We would like to express our appreciation to the City of Toronto and the Open Data Portal for granting access to the dataset via the Open Data Toronto package [@opendatatoronto].

### 2024-09-26 scripts/02-data_cleaning.R
Use code to take a review on the cleaned data.

@@ -80,7 +80,8 @@ crime_long_1 <- crime_data_sf %>%
   )) %>%
   drop_na(Crime_Count)  # Remove rows where Crime_Count is NA
 # Check the result
-head(crime_long_1)
+glimpse(crime_long_1)
+
 
 #Get the crime count for different neighbourhood in different years
 crime_summary_multiple_years <- crime_data %>%

### 2024-09-26 scripts/02-data_cleaning.R
Delete useless csv file.

@@ -36,7 +36,6 @@ crime_data_sf <- lapply(coordinates_list, function(coords) {
 })
 crime_data$geometry <- st_sfc(crime_data_sf)
 crime_data_sf <- st_as_sf(crime_data)
-
 print(st_geometry(crime_data_sf))
 
 # Get the average crime rate in the whole Toronto for the line chart
@@ -52,12 +51,13 @@ crime_long <- average_rates %>%
   mutate(Year = as.numeric(Year), 
          Crime_Type = str_to_title(Crime_Type),  
          Crime_Type = case_when( 
-           Crime_Type == "Breakenter" ~ "Break and Enters",
+           Crime_Type == "Breakenter" ~ "Break and Enter",
            Crime_Type == "Autotheft" ~ "Auto Theft",
            Crime_Type == "Biketheft" ~ "Bike Theft" ,
            Crime_Type == "Theftover" ~ "Theft Over" ,
            TRUE ~ Crime_Type
          )) %>% filter(Crime_Type != "Theftfrommv") 
+crime_long<- crime_long %>% rename(Average_Crime_Rate=value)
 head(crime_long)
 
 # Clean the dataset for the combination of maps for per type of crime
@@ -145,7 +145,7 @@ print(bottom_10_neighborhoods_consistent)
 
 #### Save data ####
 saveRDS(crime_data_sf, "data/analysis_data/cleaning_crime_data.rds")
-write_csv(crime_data_sf, "data/analysis_data/cleaning_crime_data.csv")
+#write_csv(crime_data_sf, "data/analysis_data/cleaning_crime_data.csv")
 write_csv(crime_long, "data/analysis_data/toronto_crime_average_rates.csv")
 write_rds(crime_long_1, "data/analysis_data/cleaning_crime_count.rds")
 write_csv(top_10_neighborhoods_consistent, "data/analysis_data/top_10_neighborhoods.csv")

### 2024-09-26 paper/paper.qmd
Update the paper.

@@
 ---
-title: "Exploring the Impact of Population Density on Crime Distribution in Toronto Neighborhoods (2014–2023)"
+title: "ssessing Population Density and Crime Patterns in Toronto Neighborhoods (2014–2023)"
 author: 
@@
 # Data {#sec-data}
-## Data Source
+## Data Overview
 The Neighborhood Crime Rates dataset, obtained from @opendatatoronto, spans from 2014 to 2023 and covers 158 Toronto neighborhoods.It is designed to offer communities insights into public safety and awareness.
@@
 Besides, crime counts per crime type per year are also shown in the data.
-In @opendatatoronto, another variable of interest *POPULATION_2023* refers to the population in the 2023 year in each neighborhood, enabling the analysis of the relationship between population density and crime frequency.
+In @opendatatoronto, another variable of interest *Population 2023* refers to the population in the 2023 year in each neighborhood, enabling the analysis of the relationship between population density and crime frequency.
 Geometry column is transformed into spatial data to build up the Toronto city map to observe population density and crime distribution.
+@tbl-rawdata gives us a glimpse of how the dataset looks.
+Each neighborhood is identified by its *Area Name* and a unique *Hood ID*.
 
+The average crime rates represent the mean crime rate across 158 neighborhoods from 2014 to 2023, and were calculated to analyze long-term trends and highlight crime types that have seen significant increases.
+The variables *Crime Type* and *Crime Count* are based on 2023 data (@tbl-summary1).Each neighborhood’s average rank was determined by ranking all 158 neighborhoods based on their total annual crime count.
+A rank is assigned for each year, and the average rank over the past decade (2014–2023) is calculated by averaging these yearly ranks.
+A lower rank indicates a higher crime count.
+Summaries of the top 10 and bottom 10 ranked neighborhoods are provided in @tbl-summary2 and @tbl-summary4.
+
+```{r}
+#| label: tbl-rawdata
+#| tbl-cap: A Lookup for the Raw Data
+#| echo: false
+
+crime_data <- read.csv(here::here("data/raw_data/raw_crime_data.csv"))
+crime_data <- crime_data |> 
+  select(HOOD_ID, POPULATION_2023, ASSAULT_2023, ASSAULT_RATE_2023)
+crime_data <- crime_data |> 
+  rename(
+    `Hood ID` = HOOD_ID,
+    `Population 2023` = POPULATION_2023,
+    `Assault 2023` = ASSAULT_2023,
+    `Assault Rate 2023` = ASSAULT_RATE_2023
+  )
+
+kable(head(crime_data))
+```
+
+```{r}
+#| label: tbl-summary1
+#| tbl-cap: Summary Table for the Crime Count in 2023
+#| echo: false
+#| warning: false
+#| message: false
+
+crime_count <- readRDS(here::here("data/analysis_data/cleaning_crime_count.rds"))
+crime_count_summary <- crime_count %>%
+  st_drop_geometry() %>%  
+  select(HOOD_ID, AREA_NAME, Crime_Type, Crime_Count)
+crime_count_summary <- crime_count_summary %>%
+  rename(
+    `Hood ID` = HOOD_ID,
+    `Area Name` = AREA_NAME,
+    `Crime Type` = Crime_Type,
+    `Crime Count` = Crime_Count)
+kable(head(crime_count_summary, 10))
+```
+
+```{r}
+#| label: tbl-summary2
+#| tbl-cap: Summary Table for Top10 Crime-Prone Neighborhoods (2014-2023)
+#| echo: false
+#| warning: false
+#| message: false
+
+top_10_neighborhoods <- read.csv(here::here("data/analysis_data/top_10_neighborhoods.csv"))
+top_summary<-top_10_neighborhoods%>% rename(`Neighborhood ID and Name` = Neighborhood_ID_Name, `Average Rank` = Average_Rank) %>%
+  select(`Neighborhood ID and Name`, `Average Rank`)
+kable(top_summary)
+```
+
+## Measurement
 As for the measurement, the crime rate is measured using population estimates from Environics Analytics, which is in line with the standard definition by Statistics Canada.
@@
 
-@fig-rawdata gives us a glimpse of how the dataset looks. Each neighborhood is identified by its *AREA_NAME* and a unique *HOOD_ID*.
-
-```{r}
-#| label: fig-rawdata
-#| fig-cap: A Lookup for the Raw Data
-#| echo: false
-
-crime_data <- read.csv(here::here("data/raw_data/raw_crime_data.csv"))
-crime_data <- crime_data |> select(HOOD_ID, POPULATION_2023, ASSAULT_2023, ASSAULT_RATE_2023)
-kable(head(crime_data))
-```
-
-## New Variable Constructure
-The average crime rates represent the mean crime rate across 158 neighborhoods from 2014 to 2023, and were calculated to analyze long-term trends and highlight crime types that have seen significant increases.
-The variables *Crime_Type* and *Crime_Count* are based on 2023 data (@fig-summary1).
-
-Each neighborhood’s average rank was determined by ranking all 158 neighborhoods based on their total annual crime count.
-A rank is assigned for each year, and the average rank over the past decade (2014–2023) is calculated by averaging these yearly ranks.
-A lower rank indicates a higher crime count.
-Summaries of the top 10 and bottom 10 ranked neighborhoods are provided in @fig-summary2 and @fig-summary3.
-```{r}
-#| label: fig-summary1
-#| fig-cap: Summary Table for the 2023 Crime Count
-#| echo: false
-#| warning: false
-#| message: false
-
-crime_count <- readRDS(here::here("data/analysis_data/cleaning_crime_count.rds"))
-crime_count_summary <- crime_count %>%
-  st_drop_geometry() %>%  
-  select(HOOD_ID, AREA_NAME, Crime_Type, Crime_Count)
-kable(head(crime_count_summary, 10))
-```
-
-```{r}
-#| label: fig-summary2
-#| fig-cap: Summary Table for Top10 Ranked Neighborhoods
-#| echo: false
-#| warning: false
-#| message: false
-
-top_10_neighborhoods <- read.csv(here::here("data/analysis_data/top_10_neighborhoods.csv"))
-top_summary<-top_10_neighborhoods |> select(Neighborhood_ID_Name, Average_Rank)
-kable(top_summary)
-```
-
-```{r}
-#| label: fig-summary3
-#| fig-cap: Summary Table for Bottom 10 Ranked Neighborhoods
-#| echo: false
-#| warning: false
-#| message: false
-
-bottom_10_neighborhoods<- read.csv(here::here("data/analysis_data/bottom_10_neighborhoods.csv"))
-bottom_summary<-bottom_10_neighborhoods |> select(Neighborhood_ID_Name, Average_Rank)
-kable(bottom_summary)
-```
-
 \newpage
@@
 ## Crime Rates Over the Past Decade
-@fig-rates uses the variable *Average_Crime_Rate* (@fig-summary4) to show the crime rate trend between 2014 and 2023 and demonstrate the evolution of different crime types in the city of Toronto.
+@fig-rates uses the variable *Average Crime Rate* (@tbl-summary3) to show the crime rate trend between 2014 and 2023 and demonstrate the evolution of different crime types in the city of Toronto.
 It can be seen that the visualization reveals certain patterns in some specific crime types.
@@
 ```{r}
-#| label: fig-summary4
-#| fig-cap: Average Crime Rate in Toronto
+#| label: tbl-summary3
+#| tbl-cap: Average Crime Rate in Toronto
 #| echo: false
@@
 average_crime_rate <- read.csv(here::here("data/analysis_data/toronto_crime_average_rates.csv"))
-#top_summary<-top_10_neighborhoods |> select(Neighborhood_ID_Name, Average_Rank)
+average_crime_rate <- average_crime_rate %>%
+  rename(
+    `Area Name` = AREA_NAME,
+    `Crime Type` = Crime_Type,
+    `Average Crime Rate` = Average_Crime_Rate
+  ) %>%
+  select(`Area Name`, Year, `Crime Type`, `Average Crime Rate`, )
 average_crime_rate |>
@@
 #| label: fig-rates
-#| fig-cap: Average Crime Rates in Toronto Over Time
+#| fig-cap: Average Crime Rates per 100,000 population in Toronto (2014-2023)
 #| echo: false
@@
 #| message: false
-
-crime_long <- read.csv(here::here("data/analysis_data/toronto_crime_average_rates.csv"))
+crime_long<-read.csv(here::here("data/analysis_data/toronto_crime_average_rates.csv"))
 ggplot(crime_long, aes(x = Year, y = Average_Crime_Rate, color = Crime_Type, group = Crime_Type)) +
@@
   geom_point(size = 2) +  
-  labs(title = "Average Crime Rates in Toronto Over Time",
-       x = "Year", y = "Average Crime Rate per 100,000 population",
+  labs(title = "Average Crime Rate in Toronto (2014-2023)",
+       x = "Year", y = "Average Crime Rate ",
        color = "Crime Type") +
   theme_minimal() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))  
+  theme(axis.text.x = element_text(angle = 45, hjust = 1),
+        legend.position = "bottom",  
+        plot.title = element_text(face = "bold", hjust = 0.5))
 ```
@@
 ## Toronto Population Density
-@fig-Population shows the population density in Toronto, constructed by using variable *POPULATION_2023* and geometry data.
+@fig-Population shows the population density in Toronto, constructed by using variable *Population 2023* and geometry data.
 Higher saturation represents higher population density in 2023.
@@
 #| label: fig-Population
-#| fig-cap: Population Density in Toronto
+#| fig-cap: Toronto Population Density (2023)
 #| echo: false
@@
 Each crime type is represented by a separate plot, and a linear trendline has been added to demonstrate the relationship between crime count and population size.
-
 ```{r,fig.width=8, fig.height=6}
@@
 **Break and Enter** shows a similar trend, with higher crime counts in neighborhoods with larger populations.
-
-In contrast, crimes such as **Homicide** and **Shooting** display much smaller crime counts overall, and their trendlines suggest weaker correlations with population size.
-These crimes are less frequent, and the relationship between population and crime count is less pronounced compared to property crimes.
-
 **Robbery** and **Theft Over** also show a positive relationship with population, although the correlation is not as strong as it is for Assault or Auto Theft.
@@
 
-Overall, the graph highlights that property crimes, such as **Assault**, **Auto Theft**, and **Break and Enter**, are more likely to increase with population density, whereas violent crimes like **Homicide** and **Shooting** remain relatively low regardless of neighborhood size.
-From @fig-crimeandpopulation, these scatter plots represent upward trend between Crime Count and Population is shown.
+Overall, the graph highlights that property crimes, such as **Auto Theft**, and **Break and Enter**, are more likely to increase with population density, whereas violent crimes like **Homicide** and **Shooting** remain relatively low in Toronto regardless of neighborhood size.
+
+From @fig-crimeandpopulation, most of these scatter plots represent upward trend between Crime Count and Population is shown.
 It seems that the lack of cases in **Shooting** and **Homicide** makes most dots overlap together.
-So, the map for each crime is created respectively to help us see the distribution of these 2 particular crime types.
+These crimes are less frequent, and the relationship between population and crime count is less pronounced compared to property crimes.
 
-As for the Homicide and Shooting,.............
+Consequently, extra maps are shown in @fig-homicide and @fig-shooting to display these two relatively rare crime types' occurrence in particular neighborhoods.
+The size of red dot represents for the crime count and is marked on the area they occur.
+It can be observed that most larger size dots appear in the neighborhoods with darker shading, and most of neighborhoods with lowest saturation blue does not have a reported homicide or shooting case.
 
@@
 #| label: fig-homicide
-#| fig-cap: 2023 Homicide Distribution
+#| fig-cap: Homicide Count Distribution (2023)
 #| echo: false
@@
   filter(Crime_Type == "Homicide")
-# Check the filtered data
-#head(homicide_data)
 ggplot(crime_data_sf) +
-  geom_sf(aes(fill = POPULATION_2023), color = "black", alpha = 0.8) +  # Population density map with slight transparency
-  scale_fill_gradient(low = "#d3e5ff", high = "#08306b", name = "Population") +  # Softer gradient for better visuals
-  geom_jitter(data = homicide_data, aes(geometry = geometry, size = Crime_Count),  # Use geom_jitter to avoid overlapping dots
-              color = "red", stat = "sf_coordinates", alpha = 0.6, width = 0.002, height = 0.002) +  # Add jitter and adjust transparency
-  scale_size_continuous(name = "Homicide Count", range = c(2, 10)) +  # Dot sizes for clearer representation of crime count
+  geom_sf(aes(fill = POPULATION_2023), color = "black", alpha = 0.8) + 
+  scale_fill_gradient(low = "#d3e5ff", high = "#08306b", name = "Population") +  
+  geom_jitter(data = homicide_data, aes(geometry = geometry, size = Crime_Count),  
+              color = "red", stat = "sf_coordinates", alpha = 0.6, width = 0.002, height = 0.002) + 
+  scale_size_continuous(name = "Homicide Count", range = c(2, 10)) + 
   labs(title = "Homicide Count and Population Density",
@@
     legend.position = "bottom", 
-    legend.title = element_text(face = "bold", size = 6),  
-    legend.text = element_text(size = 6)  
-  ) +
+    legend.title = element_text(face = "bold", size = 14),  
+    legend.text = element_text(size = 10)) +
   guides(size = guide_legend(override.aes = list(alpha = 1))) 
-
 ```
@@
 #| label: fig-shooting
-#| fig-cap: Shooting Count Distribution in 2023
+#| fig-cap: Shooting Count Distribution (2023)
 #| echo: false
@@
 
-# Filter for Shooting data
 shooting_data <- crime_long_1 %>%
@@
 ggplot(crime_data_sf) +
-  geom_sf(aes(fill = POPULATION_2023), color = "black", alpha = 0.9) +  # Population density map
-  scale_fill_gradient(low = "#d3e5ff", high = "#08306b", name = "Population") +  # Color gradient for population density
-  geom_jitter(data = shooting_data, aes(geometry = geometry, size = Crime_Count),  # Add jitter to avoid overlap
-              color = "red", stat = "sf_coordinates", alpha = 0.6, width = 0.002, height = 0.002) +  # Jitter points
-  scale_size_continuous(name = "Shooting Count", range = c(1, 10), breaks = c(5, 10, 15, 20)) +  # Set dot size range and breaks
+  geom_sf(aes(fill = POPULATION_2023), color = "black", alpha = 0.9) +  
+  scale_fill_gradient(low = "#d3e5ff", high = "#08306b", name = "Population") +  
+  geom_jitter(data = shooting_data, aes(geometry = geometry, size = Crime_Count),  
+              color = "red", stat = "sf_coordinates", alpha = 0.6, width = 0.002, height = 0.002) +  
+  scale_size_continuous(name = "Shooting Count", range = c(1, 10), breaks = c(5, 10, 15, 20)) +  
   labs(title = "Shooting Count and Population Density",
@@
   theme(
-    panel.grid = element_blank(),  # Remove grid
+    panel.grid = element_blank(),  
     plot.title = element_text(size = 18, face = "bold"), 
     plot.subtitle = element_text(size = 14),
-    legend.position = "bottom",  # Move legend to bottom
-    legend.title = element_text(face = "bold")
-  ) +
-  guides(size = guide_legend(override.aes = list(alpha = 1)))  # Adjust legend for point size
+    legend.position = "bottom",  
+    legend.title = element_text(face = "bold")) +
+  guides(size = guide_legend(override.aes = list(alpha = 1), position = "bottom"))
 ```
@@
 After looking at the trend of crimes in the whole Toronto, it is more important to take a look into those small neighborhoods.
-Top10 safest and Top10 Crime Prone neighborhoods are shown in the following content.
-It give police and security department to take different measures to differnt neighborhoods.Top 1 crime prone neighborhoods is West Humber-Clairville.
-Based on the map, it lcoates
+The rank here is counted based on the total crime count for each neighborhood each year.
+Higher rank implies higher crime count on average.
+Top10 safest and Top10 Crime-Prone neighborhoods are shown in the following content.
+It give police and security department to take different measures to different neighborhoods.Top 1 crime-prone neighborhoods is West Humber-Clairville.
+Based on the map, it locates upper-left corner of the city map.
+Besides, as we noted in @fig-Population, both of the most density neighborhoods 1 and 95(Annex) show in the top 10 crime-prone ranked neighborhoods.
 
@@
 #| label: fig-rank-crime
-#| fig-cap: Top10 Ranked Crime Prone Neigborhoods Based on Crime Count
+#| fig-cap: Top10 Ranked Crime-Prone Neigborhoods (2014-2023)
 #| echo: false
@@
   coord_flip() +  
-  labs(title = "Top 10 Crime Prone Neighborhoods ",
-       x = "Neighborhood", 
-       y = "Average Rank (2014-2023)") +
+  labs(title = "Top 10 Crime-Prone Neighborhoods",
+       x = "Neighborhood ID and Name", 
+       y = "Average Rank") +
   theme_minimal() +
   theme(
-    plot.title = element_text(size = 12, face = "bold"), 
+    plot.title = element_text(size = 11, face = "bold"), 
     axis.text = element_text(size = 8),  
-    axis.title = element_text(size = 12),  
+    axis.title = element_text(size = 11,face="bold"),  
     panel.grid.minor = element_blank(),  
@@
 ```{r}
-#| label: fig-rank-safe
-#| fig-cap: Top10 Ranked Safest Neigborhoods Based on Crime Count
+#| label: tbl-summary4
+#| tbl-cap: Top10 Ranked Safest Neigborhoods (2014-2023)
 #| echo: false
@@
 #| message: false
-ggplot(bottom_10_neighborhoods, aes(x = reorder(Neighborhood_ID_Name, -Average_Rank), y = Average_Rank)) +
-  geom_bar(stat = "identity", fill = "steelblue", width = 0.7) +  
-  coord_flip() + 
-  labs(title = "Top 10 Safest Neighborhoods",
-       x = "Neighborhood", 
-       y = "Average Rank(2014-2023)") +
-  theme_minimal() +
-  theme(
-    plot.title = element_text(size = 12, face = "bold"),  
-    axis.text = element_text(size = 8),  
-    axis.title = element_text(size = 12),  
-    panel.grid.minor = element_blank(),  
-    panel.grid.major.x = element_blank(),  
-    panel.grid.major.y = element_line(color = "gray", linetype = "dashed")  
-  )
+
+bottom_10_neighborhoods<- read.csv(here::here("data/analysis_data/bottom_10_neighborhoods.csv"))
+bottom_summary <- bottom_10_neighborhoods %>%
+  rename(`Neighborhood ID and Name` = Neighborhood_ID_Name, `Average Rank` = Average_Rank) %>%
+  select(`Neighborhood ID and Name`, `Average Rank`)
+kable(bottom_summary)
 ```
+
 In @fig-rank-crime, neighborhoods such as West Humber-Clairville, York University Heights, and Wellington Place have consistently ranked among the top 10 crime-prone areas over the past decade.
@@
 The increasing trend in auto theft and assault in these neighborhoods underscores the need for targeted interventions and enhanced security measures.
-In contrast,@fig-rank-safe neighborhoods like Maple Leaf and Lambton Baby Point have consistently ranked as some of the safest areas in Toronto.
+In contrast,@tbl-summary4 neighborhoods like Maple Leaf and Lambton Baby Point have consistently ranked as some of the safest areas in Toronto.
 These neighborhoods tend to have lower population densities and are more affluent, with better access to resources and community support.
@@
 The analysis revealed significant variations in crime rates over time, particularly in property-related offenses such as Auto Theft and Break and Enter.
-Assault remains the most common crime, while Homicide and Robbery rates have remained stable.
+Assault remains the most common crime in past decade, while Homicide and Robbery rates have remained stable.
 The data shows an upward trend in crime rates after 2020, particularly in Assault and Auto Theft, suggesting a link between population density and increased criminal activity, especially post-pandemic.
@@
 The sharp increase in auto theft post-2020 highlights the need for enhanced security measures, while the stable yet high assault rates point to the need for sustained law enforcement efforts in vulnerable neighborhoods.
+Spatial Distribution of Crime: Crime is not randomly distributed but tends to cluster in specific neighborhoods with particular socioeconomic characteristics.
+High concentrations of both property and violent crime are observed in the downtown core, the northeast (North York), and the northwest (Etobicoke), while lower crime rates are observed in central and west Toronto.
+The study also reveals a U-shaped spatial pattern where neighborhoods with lower socioeconomic status experience higher crime rates  .
+
 The data clearly shows a significant correlation between population density and crime frequency in Toronto neighborhoods.
@@
 ## Socioeconomic Disparities and Crime Hotspots
+
+@Wang's study also highlights the significant role of socioeconomic deprivation, residential instability, and ethnic concentration in explaining the spatial distribution of crime.
+Neighborhoods characterized by high levels of deprivation and instability tend to have higher rates of violent crime, while property crimes are more widespread but still related to similar socioeconomic factors  .
 Socioeconomic factors, alongside population density, play a significant role in shaping crime patterns.
@@
 # Acknowledgements
-All data is analyzed in R (@citeR), using packages Tidyverse (@tidyverse), Dplyr(@dplyr), Geojsonsf(@geojsonsf), Httr (@httr), Knitr(@knitr) and ggplot2 (@ggplot2), sf(@sf), gt(@gt), lintr(@lintr) in order to visualize the data and communicate information intuitively.
+All data is analyzed in R [@citeR], using packages Tidyverse [@tidyverse], Dplyr [@dplyr], Geojsonsf [@geojsonsf], Httr [@httr], Knitr [@knitr] and ggplot2 [@ggplot2], sf (@sf), gt (@gt), lintr [@lintr] in order to visualize the data and communicate information intuitively.
 
@@
 
+\newpage
 # References

### 2024-09-26 paper/paper.qmd
Update the Title.

@@
 ---
-title: "ssessing Population Density and Crime Patterns in Toronto Neighborhoods (2014–2023)"
+title: "Assessing Population Density and Crime Patterns in Toronto Neighborhoods (2014–2023)"
 author: 

### 2024-09-27 scripts/00-simulate_data.R
Update scripts for some changing variables.

@@ -4,7 +4,7 @@
 # Date: 18 Sep 2024 
 # Contact: cassieliu.liu@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: Construct the sketch about the data frame roughly
+# Pre-requisites: Construct a sketch about the data frame roughly.
 # Any other information needed? NA
 
 #### Workspace setup ####
@@ -13,72 +13,59 @@ library(tidyverse)
 library(dplyr)
 set.seed(666)
 
-#### Simulate data ####
+# Simulate Neighborhoods
 n_neighborhoods <- 158
-years <- 2014:2023
+neighborhood_ids <- 1:n_neighborhoods
+neighborhood_names <- paste("Neighborhood", neighborhood_ids)
+
+# Simulate Population Data for 2023
+population_2023 <- rnorm(n_neighborhoods, mean = 20000, sd = 5000) 
+
+# Simulate Crime Counts for 2023 (using Poisson distribution) for each neighborhood
 crime_types <- c("Assault", "Robbery", "Break and Enter", "Theft Over", 
-                 "Auto Theft", "Bike Theft", "Shooting", "Homicide")
-neighborhood_data <- data.frame(
-  HOOD_ID = 1:n_neighborhoods,
-  AREA_NAME = paste("Neighborhood", 1:n_neighborhoods)
-)
-head(neighborhood_data)
+                 "Auto Theft", "Bike Theft", "Homicide", "Shooting")
 
-# Simulate crime counts for each crime type, each year and neighborhood
-simulate_crime_counts <- function() {
-  sapply(crime_types, function(x) rpois(1, lambda = sample(50:200, 1)))  # Random counts between 50 and 200
-}
-crime_data <- expand.grid(
-  HOOD_ID = neighborhood_data$HOOD_ID,
-  Year = years
-) %>%
-  rowwise() %>%
+# Simulate crime counts for each neighborhood for each crime type
+crime_counts <- sapply(crime_types, function(crime_type) {
+  rpois(n_neighborhoods, lambda = sample(50:500, 1))  # Generate 158 Poisson-distributed counts for each crime type
+})
+
+# Convert crime counts to a data frame (each crime type has 158 rows)
+crime_data_sim <- data.frame(
+  HOOD_ID = neighborhood_ids,
+  AREA_NAME = neighborhood_names,
+  POPULATION_2023 = population_2023,
[…]

### 2024-09-27 scripts/02-data_cleaning.R

@@ -1,10 +1,10 @@
 #### Preamble ####
-# Purpose: Clean the raw data and make it available for construcing graphs
+# Purpose: Clean the raw data and make it available for visualization
 # Author: Yongqi Liu
 # Date: 20 Sep 2024 
 # Contact: cassieliu.liu@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: Have the raw data ready for processing
+# Pre-requisites: Have the raw data ready for cleaning
 
 #### Workspace setup ####
 #install.packages("sf") for handling the spatial data
@@ -15,6 +15,7 @@ library(tidyverse)
 library(dplyr)
 
 #### Clean data ####
+
 # Handling the spatial data, split the geometry column into longitude and latitude
 crime_data <- read.csv("data/raw_data/raw_crime_data.csv")
 extract_coordinates <- function(geometry_list) {
@@ -23,8 +24,8 @@ extract_coordinates <- function(geometry_list) {
   n <- length(coordinates) / 2
   longitudes <- coordinates[1:n]
   latitudes <- coordinates[(n+1):(2*n)]
-  matrix(c(longitudes, latitudes), ncol = 2, byrow = FALSE)
-}
+  matrix(c(longitudes, latitudes), ncol = 2, byrow = FALSE)}
+
 # Apply the function to each row of the geometry column and get a list of coordinate matrices
 coordinates_list <- lapply(crime_data$geometry, extract_coordinates)
 crime_data_sf <- lapply(coordinates_list, function(coords) {
@@ -61,7 +62,6 @@ crime_long<- crime_long %>% rename(Average_Crime_Rate=value)
 head(crime_long)
 
 # Clean the dataset for the combination of maps for per type of crime
-# Transform and clean crime_long_1
 crime_long_1 <- crime_data_sf %>%
   pivot_longer(cols = c(ASSAULT_2023, ROBBERY_2023, BREAKENTER_2023, THEFTOVER_2023, AUTOTHEFT_2023,
                         BIKETHEFT_2023, SHOOTING_2023, HOMICIDE_2023), 
@@ -78,12 +78,10 @@ crime_long_1 <- crime_data_sf %>%
     Crime_Type == "SHOOTING_2023" ~ "Shooting",
     TRUE ~ Crime_Type
   )) %>%
-  drop_na(Crime_Count)  # Remove rows where Crime_Count is NA
-# Check the result
+  drop_na(Crime_Count)  
 glimpse(crime_long_1)
 
-
[…]

### 2024-09-27 scripts/03-test_data.R

@@ -18,12 +18,11 @@ print(missing_values_check)
 
 # Test for missing values
 missing_values <- crime_data %>%
-  select(Assault:Homicide, Population) %>% summarise_all(~ sum(is.na(.)))
+  select(Assault:Homicide, POPULATION_2023) %>% summarise_all(~ sum(is.na(.)))
 print(missing_values)
-
 sum(is.na(crime_data))
 
-# Check the structure of the dataset to confirm data types
+# Test the structure of the dataset to confirm data types
 str(crime_data)
 
 # Test if the crime columns are numeric and the neighborhood columns are characters
@@ -34,20 +33,20 @@ print(check_data_types)
 crime_data_range_check <- crime_data %>%
   summarise(across(Assault:Homicide, list(min = min, max = max)),
             across(ends_with("Rate"), list(min = min, max = max)),
-            Population_min = min(Population), Population_max = max(Population))
+            Population_min = min(POPULATION_2023), Population_max = max(POPULATION_2023))
 print(crime_data_range_check)
 
 # Ensure that crime rates are non-negative
 crime_rate_negative_check <- crime_data %>%
   summarise(across(ends_with("Rate"), ~ sum(. < 0)))
-print(crime_rate_negative_check)
+sum(crime_rate_negative_check)
 
 # Check for duplicate neighborhood IDs
 duplicate_neighborhood_ids <- crime_data %>%
   group_by(HOOD_ID) %>%
   summarise(n = n()) %>%
   filter(n > 1)
-print(duplicate_neighborhood_ids)
+sum(duplicate_neighborhood_ids)
 
 # Check for duplicate neighborhood names
 duplicate_neighborhood_names <- crime_data %>%
@@ -56,8 +55,8 @@ duplicate_neighborhood_names <- crime_data %>%
   filter(n > 1)
 print(duplicate_neighborhood_names)
 
-
-neighborhood_geometry<-readRDS("data/raw_data/simulated_neighborhoods_geometry.rds")
+# Test for the geometry data
+neighborhood_geometry<-readRDS("data/raw_data/simulated_crime_data.rds")
 
 # Check that the latitudes and longitudes are within Toronto's expected ranges
 geometry_bounds_check <- neighborhood_geometry %>%
@@ -72,15 +71,4 @@ longitude_check <- all(st_coordinates(neighborhood_geometry)[, 1] >= -79.6 &
[…]

### 2024-09-27 scripts/00-simulate_data.R
Modify the scripts to account for changes in variable names or structure.

@@ -5,7 +5,6 @@
 # Contact: cassieliu.liu@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: Construct a sketch about the data frame roughly.
-# Any other information needed? NA
 
 #### Workspace setup ####
 library(sf) 

### 2024-09-27 scripts/02-data_cleaning.R

@@ -32,9 +32,7 @@ crime_data_sf <- lapply(coordinates_list, function(coords) {
   if (all(coords[1, ] == coords[nrow(coords), ])) {
     st_polygon(list(coords))  
   } else {
-    st_linestring(coords) 
-  }
-})
+    st_linestring(coords) }})
 crime_data$geometry <- st_sfc(crime_data_sf)
 crime_data_sf <- st_as_sf(crime_data)
 print(st_geometry(crime_data_sf))

### 2024-09-27 scripts/03-test_data.R

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Check the validity of the cleaned datasets 
 # Author: Yongqi Liu
-# Date: 21 Sep 2024 
+# Date: 20 Sep 2024 
 # Contact: cassieliu.liu@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: Obtain and have the cleaned dataset ready

### 2024-09-27 paper/paper.qmd
Update paper draft.

--- 2024-09-26 paper/paper.qmd
@@
 ---
-title: "ssessing Population Density and Crime Patterns in Toronto Neighborhoods (2014–2023)"
+title: "Assessing Population Density and Crime Patterns in Toronto Neighborhoods (2014–2023)"
 author: 
   - Yongqi Liu
-thanks: "Code and data are available at:https://github.com/Cassieliu77/Crime-Frequency-and-Population-Density.git."
-date: 26 September 2024
+thanks: "Code and data are available at:https://github.com/Cassieliu77/Crime-Frequency-and-Population-Density.git"
+date: 27 September 2024
 date-format: long
 toc: true 
-abstract: "This study examines the relationship between population density and crime distribution in Toronto neighborhoods from 2014 to 2023.
-Using data from Toronto Police Services, the paper analyzes how crime patterns vary across the city and identifies key neighborhoods where crime occurance is concentrated.By focusing on the types of crime most influenced by urban density, this study offers data-driven recommendations for targeted crime prevention strategies and policy interventions.
-The findings highlight the need for more nuanced urban planning and law enforcement efforts to address the underlying causes of crime in high-density areas."
+abstract: "This study examines the relationship between population density and crime distribution across Toronto neighborhoods from 2014 to 2023.
+The analysis reveals that high-density areas experience higher crime rates, particularly for auto theft and assault, with socioeconomic factors further influencing crime patterns.
+Based on these findings, we propose targeted interventions, such as improved public surveillance and increased community engagement, to address crime in high-risk areas.
+This research offers valuable insights for policymakers and urban planners to enhance crime prevention strategies in Toronto."
 format: pdf
@@
 
-# install.packages(c("ggplot2", "dplyr", "viridis", "sf", "gt","lintr"))
+# install.packages(c("tidyverse", "opendatatoronto", ggplot2", "knitr", "dplyr", "sf", "gt","lintr","styler"))
 library(tidyverse)
-library(httr)
 library(opendatatoronto)
@@
 library(dplyr)
-#library(viridis)
 library(sf)
@@
 library(lintr)
+library(styler)
 ```
@@
 # Introduction
-Urban crime has long been a focus for researchers and policymakers due to its profound impact on public safety and quality of life.
-In a rapidly growing metropolis like Toronto, population density plays a significant role in shaping the spatial distribution of crime.
-This study explores how population density correlates with crime frequency across Toronto neighborhoods between 2014 and 2023.
-Additionally, the research examines the influence of socioeconomic factors in high-density areas, aiming to identify neighborhoods that require targeted crime prevention strategies.
-Through this analysis, we hope to shed light on the dynamics of urban crime in Toronto and contribute to more effective policies for mitigating crime in high-density areas.
-All of the analyses in the paper are conducted in @citeR.
+Urban crime is a pressing issue for policymakers and researchers, especially in rapidly growing cities like Toronto.
+Population density plays a significant role in shaping the spatial distribution of crime, as high-density environments often create more opportunities for criminal activity.
+This study examines the correlation between population density and crime patterns in Toronto neighborhoods from 2014 to 2023, with a focus on identifying key crime hotspots.
+
+Toronto's urban landscape is diverse, with neighborhoods varying widely in terms of socioeconomic status and population density.
+Previous research has established that densely populated areas tend to experience higher crime rates, particularly for property crimes such as auto theft and break-and-enter offenses.
+Moreover, socioeconomic factors such as poverty and residential instability further contribute to crime concentration in certain neighborhoods.
+This study aims to explore these relationships by analyzing data on crime occurrences across 158 neighborhoods.
+
+By examining crime trends over a decade, this research seeks to inform policymakers about the specific challenges faced by high-density neighborhoods.
+The study's findings will guide the development of targeted crime prevention strategies, emphasizing the importance of urban planning and community engagement in reducing criminal activity.
+Through this approach, we aim to offer data-driven suggestions that contribute to safer, more resilient urban environments in Toronto.
 
@@
 ## Data Overview
-The Neighborhood Crime Rates dataset, obtained from @opendatatoronto, spans from 2014 to 2023 and covers 158 Toronto neighborhoods.It is designed to offer communities insights into public safety and awareness.
+The Neighborhood Crime Rates dataset, obtained from @opendatatoronto, spans from 2014 to 2023 and covers 158 Toronto neighborhoods.It is designed to offer communities more information about public safety and awareness.
 The data provided to the Toronto Police Service by reporting parties is preliminary and may not have undergone full verification at the time of publication.
@@
     `Crime Type` = Crime_Type,
-    `Crime Count` = Crime_Count)
+    `Crime Count` = Crime_Count
+  )
 kable(head(crime_count_summary, 10))
@@
 top_10_neighborhoods <- read.csv(here::here("data/analysis_data/top_10_neighborhoods.csv"))
-top_summary<-top_10_neighborhoods%>% rename(`Neighborhood ID and Name` = Neighborhood_ID_Name, `Average Rank` = Average_Rank) %>%
+top_summary <- top_10_neighborhoods %>%
+  rename(`Neighborhood ID and Name` = Neighborhood_ID_Name, `Average Rank` = Average_Rank) %>%
   select(`Neighborhood ID and Name`, `Average Rank`)
@@
 Also note that this dataset incorporates data for the new structure of the 158 City of Toronto neighborhoods, enabling a more comprehensive geographic analysis, and it does not contain occurrences of crime that were deemed “unfounded”.
-Fields have been updated to reflect the new structure of the 158 City of Toronto neighborhoods.Population figures only account for the resident population in each region and do not include temporary populations like commuters or business patrons.
+The term "unfounded," as defined by Statistics Canada, refers to cases where a police investigation concludes that the reported offense did not happen and was not attempted.
+Fields have also been updated to reflect the new structure of the 158 City of Toronto neighborhoods.
+Population figures only account for the resident population in each region and do not include temporary populations like commuters or business patrons.
 To protect the privacy of those involved, the locations of crime incidents have been intentionally shifted to the nearest road intersection node.
-As such, all location data should be regarded as approximate, and users are cautioned not to interpret these locations as linked to any specific address or individual.
+As a result, all location data should be regarded as approximate.
 
-Crime rates offer a more balanced comparison over time as they account for population changes in the area and give us a fairer way to see the crime trends in the past decade, which explains why a similar dataset called "Police Annual Statistical Report - Reported Crimes" is not be chosen as my dataset, which only contains the reported crime count for a certain year and a certain division, which lack a standard way to show the crime trends and spatial attribute to see the geography distribution.
+Crime rates offer a more balanced comparison over time as they account for population changes in the area and give us a fairer way to see the crime trends in the past decade, which explains why a similar dataset called "Police Annual Statistical Report - Reported Crimes" is not be chosen as my dataset, which only contains the reported crime count for a certain year and a certain division.
+It lack a standard way to show the crime trends and spatial attribute to see the geography distribution and it also includes reported crimes deemed as unfounded.
 
@@
 #| message: false
-crime_long<-read.csv(here::here("data/analysis_data/toronto_crime_average_rates.csv"))
+crime_long <- read.csv(here::here("data/analysis_data/toronto_crime_average_rates.csv"))
 ggplot(crime_long, aes(x = Year, y = Average_Crime_Rate, color = Crime_Type, group = Crime_Type)) +
@@
   geom_point(size = 2) +
-  labs(title = "Average Crime Rate in Toronto (2014-2023)",
+  labs(
+    title = "Average Crime Rate in Toronto (2014-2023)",
     x = "Year", y = "Average Crime Rate ",
-       color = "Crime Type") +
+    color = "Crime Type"
+  ) +
   theme_minimal() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1),
+  theme(
+    axis.text.x = element_text(angle = 45, hjust = 1),
     legend.position = "bottom",
-        plot.title = element_text(face = "bold", hjust = 0.5))
+    plot.title = element_text(face = "bold", hjust = 0.5)
+  )
 ```
@@
 
-```{r, fig.width=12, fig.height=10}
+```{r, fig.width=14, fig.height=12}
 #| label: fig-Population
@@
 
-crime_data_sf<- readRDS(here::here("data/analysis_data/cleaning_crime_data.rds"))
+crime_data_sf <- readRDS(here::here("data/analysis_data/cleaning_crime_data.rds"))
 ggplot(crime_data_sf) +
@@
   geom_text(aes(label = HOOD_ID, geometry = geometry),
-            stat = "sf_coordinates", size = 4, color = "white", check_overlap = TRUE) + 
-  labs(title = "Toronto Neighbourhood Population Map",
+    stat = "sf_coordinates", size = 4, color = "white", check_overlap = TRUE
+  ) +
+  labs(
+    title = "Toronto Neighbourhood Population Map",
     subtitle = "Based on 2023 Population",
-       x = "Longitude", y = "Latitude") +
+    x = "Longitude", y = "Latitude"
+  ) +
   theme_minimal() +
@@
     panel.grid = element_blank(),
-    plot.title = element_text(size = 16, face = "bold"), 
-    plot.subtitle = element_text(size = 12)  ) 
+    plot.title = element_text(size = 18, face = "bold"),
+    plot.subtitle = element_text(size = 12)
+  )
 ```
@@
 The scatterplots in @fig-crimeandpopulation illustrate the relationship between crime count and population for various crime types across Toronto neighborhoods in 2023.
-Each crime type is represented by a separate plot, and a linear trendline has been added to demonstrate the relationship between crime count and population size.
+Each crime type is represented by a separate plot, and a trendline has been added to demonstrate the relationship between crime count and population size.
 ```{r,fig.width=8, fig.height=6}
 #| label: fig-crimeandpopulation
-#| fig-cap: Crime Count with Population in 2023
+#| fig-cap: Scatterplots for Crime Count and Population in 2023
 #| echo: false
@@
   scale_x_continuous(limits = c(0, 200), breaks = seq(0, 200, by = 50)) +
-  facet_wrap(~ Crime_Type) +  
-  labs(title = "Crime Count vs. Population by Neighborhoods",
+  facet_wrap(~Crime_Type) +
+  labs(
+    title = "Relationship between Crime Count and Population",
     x = "Crime Count",
     y = "Population",
-       color = "Crime Type") +
+    color = "Crime Type"
+  ) +
   theme_minimal(base_size = 13) +
-  theme(axis.title = element_text(size = 12),  
+  theme(
+    axis.title = element_text(size = 12),
     panel.grid.major = element_line(color = "gray90"),
@@
     plot.title = element_text(face = "bold", hjust = 0.5),
-        plot.subtitle = element_text(hjust = 0.5))  
+    plot.subtitle = element_text(hjust = 0.5)
+  )
 ```
@@
   scale_fill_gradient(low = "#d3e5ff", high = "#08306b", name = "Population") +
-  geom_jitter(data = homicide_data, aes(geometry = geometry, size = Crime_Count),  
-              color = "red", stat = "sf_coordinates", alpha = 0.6, width = 0.002, height = 0.002) + 
+  geom_jitter(
+    data = homicide_data, aes(geometry = geometry, size = Crime_Count),
+    color = "red", stat = "sf_coordinates", alpha = 0.6, width = 0.002, height = 0.002
+  ) +
   scale_size_continuous(name = "Homicide Count", range = c(2, 10)) +
-  labs(title = "Homicide Count and Population Density",
+  labs(
+    title = "Homicide Count and Population Density",
     subtitle = "By Neighborhoods",
-       x = "Longitude", y = "Latitude") +
+    x = "Longitude", y = "Latitude"
+  ) +
   theme_minimal() +
@@
     legend.title = element_text(face = "bold", size = 14),
-    legend.text = element_text(size = 10)) +
+    legend.text = element_text(size = 10)
+  ) +
   guides(size = guide_legend(override.aes = list(alpha = 1)))
@@
   scale_fill_gradient(low = "#d3e5ff", high = "#08306b", name = "Population") +
-  geom_jitter(data = shooting_data, aes(geometry = geometry, size = Crime_Count),  
-              color = "red", stat = "sf_coordinates", alpha = 0.6, width = 0.002, height = 0.002) +  
+  geom_jitter(
+    data = shooting_data, aes(geometry = geometry, size = Crime_Count),
+    color = "red", stat = "sf_coordinates", alpha = 0.6, width = 0.002, height = 0.002
+  ) +
   scale_size_continuous(name = "Shooting Count", range = c(1, 10), breaks = c(5, 10, 15, 20)) +
-  labs(title = "Shooting Count and Population Density",
+  labs(
+    title = "Shooting Count and Population Density",
     subtitle = "By Neighborhood",
-       x = "Longitude", y = "Latitude") +
+    x = "Longitude", y = "Latitude"
+  ) +
   theme_minimal() +
@@
     legend.position = "bottom",
-    legend.title = element_text(face = "bold")) +
+    legend.title = element_text(face = "bold")
+  ) +
   guides(size = guide_legend(override.aes = list(alpha = 1), position = "bottom"))
@@
 ## Neighborhood Analysis
-After looking at the trend of crimes in the whole Toronto, it is more important to take a look into those small neighborhoods.
+After looking at the pattern of crimes in the whole Toronto, it is more important to take a look into those small neighborhoods.
 The rank here is counted based on the total crime count for each neighborhood each year.
 Higher rank implies higher crime count on average.
-Top10 safest and Top10 Crime-Prone neighborhoods are shown in the following content.
-It give police and security department to take different measures to different neighborhoods.Top 1 crime-prone neighborhoods is West Humber-Clairville.
-Based on the map, it locates upper-left corner of the city map.
-Besides, as we noted in @fig-Population, both of the most density neighborhoods 1 and 95(Annex) show in the top 10 crime-prone ranked neighborhoods.
+Top 10 safest and Top 10 Crime-Prone neighborhoods are drawn from the average rank in past 10 years.
+Neighborhoods like West Humber-Clairville and York University Heights ranked among the most crime-prone areas, likely due to their high population density (around 30,000 population) and busy commercial districts.
+This supports existing criminological theories that link high-density environments with increased opportunities for crime.
+Besides, as we noted in @fig-Population, both of the most density neighborhoods 1 (West Humber-Clairville) and 95 (Annex) show in the Top 10 crime-prone ranked neighborhoods.
+In addition to West Humber-Clairville and Annex, Wellington Place, Kensington-Chinatown, and Moss Park also ranked among these most crime-prone areas over the past decade.
+These neighborhoods are characterized by higher population densities and socioeconomic challenges, making them more vulnerable to crime.
+The increasing trend in auto theft and assault in these neighborhoods underscores the need for targeted interventions and enhanced security measures.
 
@@
   coord_flip() +
-  labs(title = "Top 10 Crime-Prone Neighborhoods",
+  labs(
+    title = "Top 10 Crime-Prone Neighborhoods",
     x = "Neighborhood ID and Name",
-       y = "Average Rank") +
+    y = "Average Rank"
+  ) +
   theme_minimal() +
@@
     axis.text = element_text(size = 8),
-    axis.title = element_text(size = 11,face="bold"),  
+    axis.title = element_text(size = 11, face = "bold"),
     panel.grid.minor = element_blank(),
@@
 
-bottom_10_neighborhoods<- read.csv(here::here("data/analysis_data/bottom_10_neighborhoods.csv"))
+bottom_10_neighborhoods <- read.csv(here::here("data/analysis_data/bottom_10_neighborhoods.csv"))
 bottom_summary <- bottom_10_neighborhoods %>%
@@
 
-In @fig-rank-crime, neighborhoods such as West Humber-Clairville, York University Heights, and Wellington Place have consistently ranked among the top 10 crime-prone areas over the past decade.
-These neighborhoods are characterized by higher population densities and socioeconomic challenges, making them more vulnerable to crime.
-The increasing trend in auto theft and assault in these neighborhoods underscores the need for targeted interventions and enhanced security measures.
-In contrast,@tbl-summary4 neighborhoods like Maple Leaf and Lambton Baby Point have consistently ranked as some of the safest areas in Toronto.
-These neighborhoods tend to have lower population densities and are more affluent, with better access to resources and community support.
-The stability of crime rates in these areas suggests that socioeconomic factors play a critical role in maintaining low crime levels.
-\newpage
+In contrast, @tbl-summary4 shows the neighborhoods with least crime occurrences in past 10 years.
+Neighborhoods such as Maple Leaf and Lambton Baby Point have consistently ranked as some of the safest areas in Toronto.
+These neighborhoods tend to have lower population densities, and all of them just have a population around 10,000 in 2023.
+This significant difference is likely due to their suburban residential character, lower population densities, and enhanced social cohesion.”
 
@@
 ## Summary of Findings
-The analysis revealed significant variations in crime rates over time, particularly in property-related offenses such as Auto Theft and Break and Enter.
-Assault remains the most common crime in past decade, while Homicide and Robbery rates have remained stable.
-The data shows an upward trend in crime rates after 2020, particularly in Assault and Auto Theft, suggesting a link between population density and increased criminal activity, especially post-pandemic.
+The analysis reveals notable fluctuations in crime rates over time, with significant increases in property crimes like Auto Theft and Break and Enter.
+Assault has remained the most prevalent crime over the past decade.
+A marked rise in crime rates, particularly in Assault and Auto Theft, was observed after 2020, suggesting a potential link between population density and heightened criminal activity, especially in the post-pandemic period.
+
+High-crime neighborhoods often have greater population densities, which can create more opportunities for crime due to increased anonymity and weakened social cohesion.
+These areas are frequently economic hubs or feature mixed-use development, which may elevate the risk of crimes such as theft or assault.
+
+Conversely, safer neighborhoods tend to have lower population densities and greater residential stability.
+The reduced anonymity and slower pace of life may contribute to lower crime rates.
+Additionally, these areas often benefit from stronger community engagement and better surveillance infrastructure, further deterring criminal activity.
 
@@
 The increasing trend in crimes like auto theft and assault calls for immediate attention from policymakers.
-As shown in the data, crime rates do not arise uniformly across all neighborhoods but are concentrated in specific areas.
-This concentration suggests that socioeconomic factors, along with population density, play a significant role in the prevalence of crime.
 The sharp increase in auto theft post-2020 highlights the need for enhanced security measures, while the stable yet high assault rates point to the need for sustained law enforcement efforts in vulnerable neighborhoods.
-Spatial Distribution of Crime: Crime is not randomly distributed but tends to cluster in specific neighborhoods with particular socioeconomic characteristics.
+
+Crime is not randomly distributed but tends to cluster in specific neighborhoods with particular socioeconomic characteristics.
 High concentrations of both property and violent crime are observed in the downtown core, the northeast (North York), and the northwest (Etobicoke), while lower crime rates are observed in central and west Toronto.
-The study also reveals a U-shaped spatial pattern where neighborhoods with lower socioeconomic status experience higher crime rates  .
+Neighborhoods with high levels of marginalization tend to see higher crime rates, reinforcing theories of social disorganization and relative deprivation [@Wang].
+The findings also support criminological theories that increased population density creates more opportunities for crime due to social fragmentation and anonymity.
 
-The data clearly shows a significant correlation between population density and crime frequency in Toronto neighborhoods.
-Densely populated areas such as Moss Park, St.
-Lawrence-East Bayfront, and Kensington-Chinatown consistently rank among the highest in terms of crime rates.
-The findings support criminological theories suggesting that increased population density creates more opportunities for crime due to social fragmentation and anonymity.
 Urban planners and policymakers must consider interventions such as improved public surveillance and community engagement initiatives to reduce crime in high-density areas.
-The relationship between population density and crime suggests that urban planners and local governments should consider designing interventions that specifically address high-density environments.
-Strategies such as better lighting, increased public surveillance (e.g., CCTV), and more frequent police patrols can act as deterrents in areas where crime tends to cluster.
-Additionally, promoting community cohesion through events, neighborhood associations, and resident engagement initiatives can reduce the social disorganization that often accompanies densely populated urban settings.
+They should focus on strategies designed specifically for high-density areas.
+Solutions such as improved public lighting, increased surveillance by CCTV and more frequent police patrols can help deter criminal activity in these hotspots.
+Furthermore, fostering community engagement through neighborhood associations, local events, and resident-driven initiatives can reduce social disorganization and strengthen neighborhood resilience.
 
 ## Socioeconomic Disparities and Crime Hotspots
+@Wang's study highlights the significant role of socioeconomic deprivation, residential instability, and ethnic concentration in explaining the spatial distribution of crime.
+Neighborhoods marked by high levels of deprivation and instability tend to experience more violent crime, while property crimes, though more widespread, are still linked to similar socioeconomic factors.
 
-@Wang's study also highlights the significant role of socioeconomic deprivation, residential instability, and ethnic concentration in explaining the spatial distribution of crime.
-Neighborhoods characterized by high levels of deprivation and instability tend to have higher rates of violent crime, while property crimes are more widespread but still related to similar socioeconomic factors  .
-Socioeconomic factors, alongside population density, play a significant role in shaping crime patterns.
-Neighborhoods with higher levels of economic disadvantage, such as Moss Park and York University Heights, tend to have elevated crime rates, while affluent areas like Kingsway South report significantly lower crime levels.
-This suggests that addressing socioeconomic disparities is essential for reducing crime.
-Targeted social interventions, such as improving access to education, healthcare, and employment opportunities, could help mitigate the underlying causes of crime.
-@fig-rates shows several crime types, notably Assault and Auto Theft, experienced noticeable increases after 2020, possibly indicating societal shifts, such as the impacts of the pandemic on public safety and economic conditions.
-Crime rates also vary between urban and rural areas.
-Densely populated urban areas typically experience higher crime rates due to greater social anonymity, economic disparities, and more opportunities for criminal activity.
-Conversely, remote rural areas often face challenges related to limited police resources, complicating crime prevention and law enforcement efforts.
-Socioeconomic inequality drives crime dynamics by creating social strain and increasing the likelihood of economic crimes such as theft, robbery, and property damage.
-Policymakers should prioritize reducing economic inequality through programs that address poverty, unemployment, and lack of access to education.
-Neighborhood revitalization initiatives that improve housing, create job opportunities, and enhance access to essential services like education and healthcare could help address the root causes of crime and reduce its prevalence in disadvantaged areas.
+Socioeconomic factors, together with population density, play a key role in shaping crime patterns.
+Economically disadvantaged areas, such as Moss Park and York University Heights, tend to have higher crime rates.
+This underscores the importance of addressing socioeconomic disparities to reduce crime.
+Targeted social interventions—such as improving access to education, healthcare, and employment—could help tackle the root causes of crime.
+
+@mohammadi's research reinforces this view, suggesting that neighborhoods with high levels of marginalization often see elevated crime rates, supporting theories of social disorganization and relative deprivation.
+The presence of commercial establishments and large buildings is also positively associated with homicide rates, likely due to increased social interactions and conflicts in these busier, more economically active areas.
+
+Socioeconomic inequality fuels crime dynamics by creating social strain and increasing the likelihood of economic crimes like theft, robbery, and property damage.
+Policymakers should focus on reducing economic inequality through programs that target poverty, unemployment, and lack of access to education.
+Revitalizing disadvantaged neighborhoods by improving housing, creating job opportunities, and enhancing access to essential services like education and healthcare could help address the underlying causes of crime and reduce its incidence in these areas.
 
 ##  Targeted Crime Prevention Strategies
-@Wang's spatial analyses support previous scholarly findings that crime is not randomly dispersed across areas, but is instead concentrated in neighborhoods that exhibit specific characteristics.
-The geographic concentration of crime in specific neighborhoods has significant implications for law enforcement and public policy.
-Policymakers and law enforcement agencies must recognize that crime prevention cannot be approached with a one-size-fits-all strategy.
-Instead, targeted interventions are necessary to address the unique challenges faced by high-risk neighborhoods.
-he data reinforces the idea that crime is not uniformly distributed across Toronto neighborhoods but is concentrated in specific high-risk areas.
-This suggests that a “one-size-fits-all” approach to crime prevention would be ineffective.
-Instead, law enforcement agencies should focus their efforts on areas where crime is most prevalent, such as Moss Park and Kensington-Chinatown, by deploying more officers and allocating more resources to these neighborhoods.
+A pattern has been verified in Toronto is that persistently high crime rates often draw additional criminal activity, leading to what is known as the “law of crime concentrations at places” [@mohammadi].
+It means crimes tend to recur in some areas with specific features, thus these features can be found out to take crime prevention strategies.
+Property and violent crime hotspots are concentrated in the downtown core, northwest, and eastern parts of the city [@Wang].
+This geographic concentration has significant implications for law enforcement and public policy.
+Policymakers and law enforcement agencies must recognize that crime prevention strategies cannot be one-size-fits-all.
+Instead, interventions need to be tailored to address the unique challenges faced by high-risk neighborhoods.
 
-At the same time, the study highlights the importance of community-based crime prevention strategies.
-Engaging local residents in neighborhood watch programs, providing education on crime prevention techniques, and fostering collaboration between police and community organizations can empower communities to take ownership of their safety.
-This approach not only reduces crime rates but also builds trust between law enforcement and the communities they serve, which is essential for long-term success in crime reduction.
+This study underscores the importance of community-based crime prevention strategies.
+Engaging local residents through neighborhood watch programs, educating the public on crime prevention techniques, and fostering collaboration between the police and community organizations can empower communities to take charge of their safety.
+Such an approach not only helps reduce crime frequency but also strengthens the relationship between law enforcement and the communities they serve—an essential element for long-term success in crime reduction.
 
+@Wang propose that neighborhoods with higher levels of socioeconomic disadvantage—marked by unemployment, poverty, and lower education levels tend to experience higher crime rates.
+This aligns with social disorganization theory, which links socioeconomic disadvantages to an increase in crime.
 For neighborhoods like Moss Park and St.
-Lawrence-East Bayfront, which consistently rank among the highest neighborhodos in crimes, a multifaceted approach is required.
-This might include increased policing and community-based programs aimed at addressing the root causes of crime, such as poverty, unemployment, and lack of access to education and healthcare.
-Research suggests that addressing these underlying issues can have a significant impact on reducing crime rates, particularly in economically disadvantaged areas.
+Lawrence-East Bayfront, which consistently rank among the areas with the highest crime rates, a multifaceted approach is necessary.
+This could include increased policing along with community programs that address the root causes of crime, such as poverty, unemployment, and lack of access to education and healthcare.
+Tackling these underlying issues can significantly reduce crime rates, particularly in economically disadvantaged areas.
 
-Furthermore, the rise in auto theft and break-and-enter incidents post-2020 suggests that law enforcement must adapt to changing crime patterns.
-Technological advancements, such as the increased use of vehicle tracking systems and smart security devices, could help curb the rising rates of property crime.
-In addition, public awareness campaigns aimed at promoting better security practices, particularly in high-risk neighborhoods, may also prove effective in preventing crime.
+Additionally, the rise in auto theft and break-and-enter incidents in recent years indicates that law enforcement must adapt to evolving crime patterns.
+The information era has introduced new types of crime, such as cybercrime and telecom crime.
+These types of crimes are missing from our dataset.
+Technological advancements, like vehicle tracking systems and smart security devices, can help curb these rising property crime rates.
+Public awareness campaigns promoting better security practices, especially in high-risk neighborhoods, could also be effective in preventing crime.
 
 ##  Limitations and Future Research
-While this study represents the relationship between population density and crime frequency in Toronto, several limitations must be acknowledged.
-Firstly, the analysis relies on reported crime data, which may not fully capture the extent of criminal activity, particularly for underreported crimes such as domestic violence and cybercrime.
-@Oliveira's findings indicate that the assumption of linear crime growth is inaccurate.
-In over half of the analyzed data sets, we observed evidence of nonlinear crime growth, meaning that crime tends to increase with population size at a rate different from per capita expectations.
-This nonlinearity introduces a population effect that impacts crime rates and alters city rankings.
-We showed that ranking cities based on crime rates alone yields significantly different results compared to rankings that account for population size.
-So, the further
+While this study provides a strong foundation for assessing population density and crime patterns in Toronto neighborhoods, some limitations must be acknowledged.
+Firstly, the analysis relies on reported crime data, which may not fully reflect the actual scope of criminal activity, especially for underreported crimes like domestic violence and cybercrime.
+Information era has led to some new types of criminal activities like cybercrime and telecom crime.
+Obtaining a definitive national cybercrime rate or count is still a challenging task because of underreporting.
 
-Secondly, the study focuses primarily on the relationship between population density and crime rates, without accounting for other factors that may also influence crime, such as the presence of law enforcement, urban infrastructure, and even social events.
+Secondly, this study primarily focuses on the relationship between population density and crime patterns, without considering other factors that may also influence the crime data, such as law enforcement presence, urban infrastructure, or social events.
 Future research could incorporate these variables to provide a more nuanced understanding of crime dynamics in urban areas.
 
-Finally, the analysis spans a relatively short period (2014–2023), which may not capture long-term trends or shifts in crime patterns.
-Expanding the dataset to include earlier years or incorporating data from other cities could offer a broader perspective on crime dynamics in Toronto and beyond.
-
-Besides, @Wang mentions the digital era has ushered in a new type of criminal activity: cybercrime and telecom crime.
-The Canadian Centre for Cyber Security (CCCS) reports a consistent rise in cybercrime incidents, which amounted to 74,073 police-reported cases as of recent data from Statista @Wang.
-This kind of crime count is missing in our dataset.While obtaining a definitive national cybercrime rate is challenging due to underreporting, the CCCS highlights phishing scams, malware attacks, and ransomware as significant threats.
-Besides, as for differnt crime types, there may be differnt potential factors influecning the crime occurance.
-
-Future research should expand on these findings by incorporating additional data sources to capture a broader range of criminal activity.
-It would also be useful to explore how different types of urban infrastructure, such as public transportation or lighting, influence crime rates.
-Finally, comparative studies with other major cities could provide a more comprehensive understanding of urban crime dynamics, enabling better-targeted interventions for crime prevention.
+Future research could expand on these findings by incorporating additional data sources and other potential influences to capture a wider range of criminal activity.
+It would also be valuable to explore how different aspects of urban infrastructure and development of new technology affect crime patterns.
 
@@
 # Acknowledgements
-All data is analyzed in R [@citeR], using packages Tidyverse [@tidyverse], Dplyr [@dplyr], Geojsonsf [@geojsonsf], Httr [@httr], Knitr [@knitr] and ggplot2 [@ggplot2], sf (@sf), gt (@gt), lintr [@lintr] in order to visualize the data and communicate information intuitively.
+All data analysis was performed in R [@citeR], utilizing a range of packages including Tidyverse [@tidyverse], Dplyr [@dplyr], Geojsonsf [@geojsonsf], Knitr [@knitr], ggplot2 [@ggplot2], sf [@sf], gt [@gt], and lintr [@lintr].
+These tools facilitated both data visualization and the clear communication of findings.
 
-We would like to express our appreciation to the City of Toronto and the Open Data Portal for granting access to the dataset via the Open Data Toronto package [@opendatatoronto].
-Besides, R codes from @citerohan are used to provide tables in this paper.The provision of high-quality, open-source data is essential for enabling precise research.
-Thanks to @openai2023chatgpt, which is used in this research paper.Code written in the scripts was checked and styled with lintr [@citelintr] and styler [@citestyler].
+We extend our gratitude to the City of Toronto and the Open Data Portal for providing access to the dataset via the Open Data Toronto package [@opendatatoronto].
+Additionally, R code from @rohan was utilized to generate tables in this paper.
+We also acknowledge the use of ChatGPT by OpenAI [@openai2023chatgpt].
+Finally, the code was reviewed and formatted using lintr [@lintr] and styler [@styler].
 

### 2024-09-27 paper/paper.qmd
Revise small typos.

@@
 toc: true 
-abstract: "This study examines the relationship between population density and crime distribution across Toronto neighborhoods from 2014 to 2023.
+abstract: "This paper examines the relationship between population density and crime distribution across Toronto neighborhoods from 2014 to 2023.
 The analysis reveals that high-density areas experience higher crime rates, particularly for auto theft and assault, with socioeconomic factors further influencing crime patterns.
-Based on these findings, we propose targeted interventions, such as improved public surveillance and increased community engagement, to address crime in high-risk areas.
-This research offers valuable insights for policymakers and urban planners to enhance crime prevention strategies in Toronto."
+Based on these findings, we propose targeted interventions and offers data-driven suggestions for policymakers and urban planners to enhance crime prevention strategies in Toronto."
 format: pdf
@@
 A lower rank indicates a higher crime count.
-Summaries of the top 10 and bottom 10 ranked neighborhoods are provided in @tbl-summary2 and @tbl-summary4.
+Summaries of the Top 10 and Bottom 10 ranked neighborhoods are provided in @tbl-summary2 and @tbl-summary4.
 
@@
 #| label: tbl-summary2
-#| tbl-cap: Summary Table for Top10 Crime-Prone Neighborhoods (2014-2023)
+#| tbl-cap: Summary Table for Top 10 Crime-Prone Neighborhoods (2014-2023)
 #| echo: false
@@
 
-\newpage
 # Results
@@
 The crime rate for **Assault** consistently remained the highest across all crime types, starting at around 600 per 100,000 population in 2014 and reaching a peak of approximately 800 per 100,000 population in 2023.
-This consistent rise suggests that assault has been a persistent issue in Toronto, with an even sharper increase following the onset of the COVID-19 pandemic.**Break and Enter**,**Robbery** and **Theft Over** all experienced increases after 2021.
+This consistent rise suggests that assault has been a persistent issue in Toronto, with an even sharper increase following the onset of the COVID-19 pandemic.
+**Break and Enter**,**Robbery** and **Theft Over** all experienced increases after 2021.
 This rise suggests a sharp increase in private property related crimes, possibly influenced by socioeconomic factors or changes in law enforcement or technology.
@@
     plot.title = element_text(size = 18, face = "bold"),
-    plot.subtitle = element_text(size = 12)
+    plot.subtitle = element_text(size = 13)
   )
@@
 #| label: fig-rank-crime
-#| fig-cap: Top10 Ranked Crime-Prone Neigborhoods (2014-2023)
+#| fig-cap: Top 10 Ranked Crime-Prone Neigborhoods (2014-2023)
 #| echo: false
@@
 #| label: tbl-summary4
-#| tbl-cap: Top10 Ranked Safest Neigborhoods (2014-2023)
+#| tbl-cap: Top 10 Ranked Safest Neigborhoods (2014-2023)
 #| echo: false

### 2024-09-27 paper/paper.qmd
Update github link in the footnote.

@@
   - Yongqi Liu
-thanks: "Code and data are available at:https://github.com/Cassieliu77/Crime-Frequency-and-Population-Density.git"
+thanks: "Code and data are available at:https://github.com/Cassieliu77/Crime-Patterns-and-Population-Density.git"
 date: 27 September 2024

### 2024-09-27 paper/paper.qmd
Revise some typos.

@@
 The analysis reveals that high-density areas experience higher crime rates, particularly for auto theft and assault, with socioeconomic factors further influencing crime patterns.
-Based on these findings, we propose targeted interventions and offers data-driven suggestions for policymakers and urban planners to enhance crime prevention strategies in Toronto."
+Based on these findings, we propose targeted interventions and offer data-driven suggestions for policymakers and urban planners to enhance crime prevention strategies in Toronto."
 format: pdf
@@
 ## Data Overview
-The Neighborhood Crime Rates dataset, obtained from @opendatatoronto, spans from 2014 to 2023 and covers 158 Toronto neighborhoods.It is designed to offer communities more information about public safety and awareness.
+The Neighborhood Crime Rates dataset, obtained from @opendatatoronto, spans from 2014 to 2023 and covers 158 Toronto neighborhoods.
+It is designed to offer communities more information about public safety and awareness.
 The data provided to the Toronto Police Service by reporting parties is preliminary and may not have undergone full verification at the time of publication.
@@
 In @opendatatoronto, another variable of interest *Population 2023* refers to the population in the 2023 year in each neighborhood, enabling the analysis of the relationship between population density and crime frequency.
-Geometry column is transformed into spatial data to build up the Toronto city map to observe population density and crime distribution.
+The geometry column is transformed into spatial data to build up the Toronto city map to observe population density and crime distribution.
 @tbl-rawdata gives us a glimpse of how the dataset looks.
@@
 
-The average crime rates represent the mean crime rate across 158 neighborhoods from 2014 to 2023, and were calculated to analyze long-term trends and highlight crime types that have seen significant increases.
-The variables *Crime Type* and *Crime Count* are based on 2023 data (@tbl-summary1).Each neighborhood’s average rank was determined by ranking all 158 neighborhoods based on their total annual crime count.
+The average crime rates represent the mean crime rate across 158 neighborhoods from 2014 to 2023 and were calculated to analyze long-term trends and highlight crime types that have seen significant increases.
+The variables *Crime Type* and *Crime Count* are based on 2023 data (@tbl-summary1).
+Each neighborhood’s average rank was determined by ranking all 158 neighborhoods based on their total annual crime count.
 A rank is assigned for each year, and the average rank over the past decade (2014–2023) is calculated by averaging these yearly ranks.
@@
 Crime rates offer a more balanced comparison over time as they account for population changes in the area and give us a fairer way to see the crime trends in the past decade, which explains why a similar dataset called "Police Annual Statistical Report - Reported Crimes" is not be chosen as my dataset, which only contains the reported crime count for a certain year and a certain division.
-It lack a standard way to show the crime trends and spatial attribute to see the geography distribution and it also includes reported crimes deemed as unfounded.
+It lacks a standard way to show the crime trends and spatial attributes to see the geography distribution and it also includes reported crimes deemed as unfounded.
 
@@
 **Break and Enter**,**Robbery** and **Theft Over** all experienced increases after 2021.
-This rise suggests a sharp increase in private property related crimes, possibly influenced by socioeconomic factors or changes in law enforcement or technology.
+This rise suggests a sharp increase in private property-related crimes, possibly influenced by socioeconomic factors or changes in law enforcement or technology.
 In contrast, **Homicide** and **Shooting** had relatively lower crime rates and displayed more stability over time, with little significant change compared to **Assault** and **Auto Theft**.
@@
 Higher saturation represents higher population density in 2023.
-Neighborhoods with lighter shading have smaller populations, such as neighborhood 12 and 13.
+Neighborhoods with lighter shading have smaller populations, such as neighborhoods 12 and 13.
 Notably, neighborhoods 1 and 95 exhibit the darkest saturation, indicating the highest population density.
@@
 
-From @fig-crimeandpopulation, most of these scatter plots represent upward trend between Crime Count and Population is shown.
+From @fig-crimeandpopulation, most of these scatter plots represent an upward trend between Crime Count and Population is shown.
 It seems that the lack of cases in **Shooting** and **Homicide** makes most dots overlap together.
@@
 Consequently, extra maps are shown in @fig-homicide and @fig-shooting to display these two relatively rare crime types' occurrence in particular neighborhoods.
-The size of red dot represents for the crime count and is marked on the area they occur.
-It can be observed that most larger size dots appear in the neighborhoods with darker shading, and most of neighborhoods with lowest saturation blue does not have a reported homicide or shooting case.
+The size of the red dot represents the crime count and is marked on the area they occur.
+It can be observed that most larger-sized dots appear in the neighborhoods with darker shading, and most of the neighborhoods with the lowest saturation blue do not have a reported homicide or shooting case.
 
@@
 ## Neighborhood Analysis
-After looking at the pattern of crimes in the whole Toronto, it is more important to take a look into those small neighborhoods.
+After looking at the pattern of crimes in the whole of Toronto, it is more important to take a look into those small neighborhoods.
 The rank here is counted based on the total crime count for each neighborhood each year.
-Higher rank implies higher crime count on average.
-Top 10 safest and Top 10 Crime-Prone neighborhoods are drawn from the average rank in past 10 years.
+A higher rank implies a higher crime count on average.
+The Top 10 safest and Top 10 Crime-Prone neighborhoods are drawn from the average rank in the past 10 years.
 Neighborhoods like West Humber-Clairville and York University Heights ranked among the most crime-prone areas, likely due to their high population density (around 30,000 population) and busy commercial districts.
 This supports existing criminological theories that link high-density environments with increased opportunities for crime.
-Besides, as we noted in @fig-Population, both of the most density neighborhoods 1 (West Humber-Clairville) and 95 (Annex) show in the Top 10 crime-prone ranked neighborhoods.
+Besides, as we noted in @fig-Population, both of the density neighborhoods 1 (West Humber-Clairville) and 95 (Annex) show in the Top 10 crime-prone ranked neighborhoods.
 In addition to West Humber-Clairville and Annex, Wellington Place, Kensington-Chinatown, and Moss Park also ranked among these most crime-prone areas over the past decade.
@@
 
-In contrast, @tbl-summary4 shows the neighborhoods with least crime occurrences in past 10 years.
+In contrast, @tbl-summary4 shows the neighborhoods with the least crime occurrences in the past 10 years.
 Neighborhoods such as Maple Leaf and Lambton Baby Point have consistently ranked as some of the safest areas in Toronto.
-These neighborhoods tend to have lower population densities, and all of them just have a population around 10,000 in 2023.
-This significant difference is likely due to their suburban residential character, lower population densities, and enhanced social cohesion.”
+These neighborhoods tend to have lower population densities, and all of them just have a population of around 10,000 in 2023.
+This significant difference is likely due to their suburban residential character, lower population densities, and enhanced social cohesion.
 
@@
 ##  Targeted Crime Prevention Strategies
-A pattern has been verified in Toronto is that persistently high crime rates often draw additional criminal activity, leading to what is known as the “law of crime concentrations at places” [@mohammadi].
-It means crimes tend to recur in some areas with specific features, thus these features can be found out to take crime prevention strategies.
+A pattern that has been verified in Toronto is that persistently high crime rates often draw additional criminal activity, leading to what is known as the “law of crime concentrations at places” [@mohammadi].
+It means crimes tend to recur in some areas with specific features, thus these features can be found to take crime prevention strategies.
 Property and violent crime hotspots are concentrated in the downtown core, northwest, and eastern parts of the city [@Wang].
@@
 This study underscores the importance of community-based crime prevention strategies.
-Engaging local residents through neighborhood watch programs, educating the public on crime prevention techniques, and fostering collaboration between the police and community organizations can empower communities to take charge of their safety.
+Engaging residents through neighborhood watch programs, educating the public on crime prevention techniques, and fostering collaboration between the police and community organizations can empower communities to take charge of their safety.
 Such an approach not only helps reduce crime frequency but also strengthens the relationship between law enforcement and the communities they serve—an essential element for long-term success in crime reduction.
@@
 Firstly, the analysis relies on reported crime data, which may not fully reflect the actual scope of criminal activity, especially for underreported crimes like domestic violence and cybercrime.
-Information era has led to some new types of criminal activities like cybercrime and telecom crime.
+The information era has led to some new types of criminal activities like cybercrime and telecom crime.
 Obtaining a definitive national cybercrime rate or count is still a challenging task because of underreporting.
@@
 Future research could expand on these findings by incorporating additional data sources and other potential influences to capture a wider range of criminal activity.
-It would also be valuable to explore how different aspects of urban infrastructure and development of new technology affect crime patterns.
+It would also be valuable to explore how different aspects of urban infrastructure and the development of new criminal means affect crime patterns.
 

### 2024-09-27 paper/paper.qmd
Revise typos.
Revise some typos.

@@
 toc: true 
-abstract: "This paper examines the relationship between population density and crime distribution across Toronto neighborhoods from 2014 to 2023.
+abstract: "This paper examines the relationship between population density and crime patterns across Toronto neighborhoods from 2014 to 2023.
 The analysis reveals that high-density areas experience higher crime rates, particularly for auto theft and assault, with socioeconomic factors further influencing crime patterns.
@@
 Through this approach, we aim to offer data-driven suggestions that contribute to safer, more resilient urban environments in Toronto.
+
+In the sections that follow, this paper outlines the data used and measurement methods, presents the results derived from the data and discusses the implications of these results.
+These findings underscore the critical need for nuanced urban planning and tailored law enforcement strategies to mitigate crime in densely populated areas of Toronto.
 
@@
     title = "Shooting Count and Population Density",
-    subtitle = "By Neighborhood",
+    subtitle = "By Neighborhoods",
     x = "Longitude", y = "Latitude"
@@
 This underscores the importance of addressing socioeconomic disparities to reduce crime.
-Targeted social interventions—such as improving access to education, healthcare, and employment—could help tackle the root causes of crime.
+Targeted social interventions such as improving access to education, healthcare, and employment could help tackle the root causes of crime.
 
